---
title: Alcohol and sexual consent communication
subtitle: A daily diary investigation of verbal and nonverbal cues
author:
  - name: Stephanie B. Ward 
    orcid: https://orcid.org/0000-0002-8191-0158
    corresponding: true
    email: sbward@wisc.edu
    affiliations:
      - Department of Psychology, University of Wisconsin-Madison
keywords:
  - Sexual consent
  - Alcohol 
  - Sexual assault
  - College student
date: last-modified
bibliography: zotero_lib.bib
csl: _csl/american-medical-association.csl
number-sections: false 
editor_options: 
  chunk_output_type: console
editor:
  wrap: 72
---

## Introduction

Sexual consent is a complex process involving thoughts, feelings, and
behaviors used to determine and express willingness to engage in sexual
activity.[@muehlenhardComplexitiesSexualConsent2016] Specific behaviors
used to infer and communicate consent include verbal and nonverbal cues,
which can be explicit or implicit, ranging from clear spoken dialogue
("Yes!") to subtle body
language.[@willisAssociationsInternalExternal2019] The presence of
consent distinguishes wanted sexual experiences from sexual assault, so
programs aiming to prevent sexual assault often target consent
communication. A substantial gap exists between these programs'
curricula and the varied consent practices that young people use during
sexual encounters, particularly those involving alcohol
use.[@curtis2017]

Alcohol use frequently co-occurs with sexual assault and muddles
communication dynamics within young adults' consensual sexual
encounters. A recent scoping review of the alcohol and sexual consent
literature highlighted the need for more quantitative research in this
area. The authors called for studies using methods that permit careful
evaluation of the contextual features of alcohol-involved consent by
facilitating event-level data collection on individual sexual encounters
and aspects within them that covary with cue use. This call to action is
important because contextual factors—such as co-occurring intoxication
and type of sexual activity—have been linked to within-person
differences in consent communication. [@rinioloRoleAlcoholContext2025,
@willisInternalExternalSexual2021]

Researchers studying contextual factors linking alcohol to variance in
consent communication have done so by conducting systematic,
within-person comparisons of two matched events,
[@marcantonioWithinPersonExaminationAlcoholInvolved2024] or by grounding
responses to each person’s most recent sexual encounter, which could
have occurred weeks to months in the past. While informative, these
studies risk greater recall bias by asking people to remember past
events. In contrast, micro-longitudinal studies assess contextual
factors closer in time to when consent communication occurs. Although
ecological momentary assessment and electronic daily diaries have been
used extensively in other clinical science research to illuminate
important time-varying predictors of health risk behaviors,
[@serreEcologicalMomentaryAssessment2015, @rodríguez-blanco2018,
@bolger2003] only one published investigation of this kind has been
examined alcohol and sexual consent communication.

Willis and colleagues surveyed participants three times per day for 28
days and found that within-person variability accounted for more than
70% of the variance in sexual consent
communication.[@willisAssessingWithinPersonVariability2021] Participants
reported greater consent cue use for sexual encounters involving more
physically intimate sexual behaviors (i.e., vaginal or anal penetration)
than for encounters involving less intimate behaviors,
[@willisAssessingWithinPersonVariability2021] but no significant
event-level associations between alcohol involvement and verbal or
nonverbal consent were observed in this
sample.[@willisInternalExternalSexual2021] The previous research
examined a limited set of predictors and reflected a specific
subpopulation of adults in committed relationships; the current project
builds on their findings by probing encounter- and person-level
associations with alcohol and consent cue use among university students
while expanding the range of possible predictors. Examining these
associations in different samples and in relation to a wider array of
variables is a crucial next step toward isolating more actionable
targets for sexual assault prevention efforts focused on changing
consent behavior.

### Between-person predictors of active consent cues

Rates of sexual assault on U.S. college campuses remain virtually
unchanged after more than 30 years of dedicated research and
intervention. [@kossScopeRapeIncidence1987,
@kossScopeRapeVictimization2022] Lack of consent is increasingly
recognized as defining characteristic of sexual assault, leading
institutions to incorporate consent-focused education into violence
prevention efforts. [@beresPerspectivesRapepreventionEducators2020,
@beresRethinkingConceptConsent2014] University policy and educational
initiatives promote explicit and enthusiastic consent communication
(“only yes means yes!”), while young people often rely instead on
passive cues (not resisting, not saying no) and assumptions (“We love
each other”) [@willisSexualConsentCommitted2021] to determine consent.
[@marcantonioEffectsTypicalBinge2022,
@jozkowskiGenderDifferencesHeterosexual2014] Sexual consent education
aims to promote the use of active cues by enhancing consent knowledge
and positive attitudes while decreasing rape-supportive attitudes,
presuming that attitudes shape consent behavior. However, these
interventions do not have a clear underlying mechanism of behavior
change, and they rarely assess behavioral outcomes.
[@beresPerspectivesRapepreventionEducators2020,
@burtonTeachingSexualConsent2023] A prerequisite to improving upon the
value of consent education is testing whether these attitudinal
intervention targets predict actual in-the-moment cue use, and
contrasting their explanatory power with that of other predictors.

### Attitudes and adherence to social norms

#### Individual differences

Rape-supportive attitudes and beliefs have been theorized to explain
variability in how individuals approach consent communication.
[@jozkowskiChapter9Sexual2022] Perhaps the most prominent example in the
sexual assault and consent literatures concerns Rape Myth Acceptance
(RMA), the endorsement of widely-held false beliefs that justify sexual
aggression and place blame on victims.
[@lonswayAttitudinalAntecedentsRape1995, @payneRapeMythAcceptance1999]
RMA is a robust predictor of sexual violence perpetration; it also has
been used as an outcome measure in several sexual consent interventions
delivered to young people in educational settings.
[@burtonTeachingSexualConsent2023, @baker2014] Individuals with high
levels of RMA often hold the belief that victims of sexual assault “led
the other person on” (e.g., with what could be conceived through this
lens as nonverbal consent cues) or provided verbal agreement and “later
regretted it.” These pervasive schemas suggest individuals with high
levels of RMA may pursue sexual contact without first communicating
their willingness or looking for indications of the other person’s
openness to their advances. [@silver2018,
@mcmahonUpdatedMeasureAssessing2011] There also are rape myths that
specifically excuse perpetrator behavior in the context of alcohol
intoxication (e.g., “It shouldn't be considered rape if the person is
drunk and didn't know what they were doing”). Rape myths are distinct
from but related to permissive Alcohol and Sexual Consent (ASC)
attitudes, [@mcmahonUpdatedMeasureAssessing2011] both of which
contradict sexual assault programming messages about, for example, how
communication signals are easily misinterpreted when alcohol is involved
in a sexual situation. Permissive ASC attitudes reflect social norms
that are prevalent on college campuses (e.g., “Consensual drunk sex is a
normal and harmless part of college life”) and reinforce approval of
intoxicated sexual activity without regard for the associated risks. Our
previous research found that permissive ASC attitudes were negatively
related to ongoing and communicative (i.e., “process-based”) consent
behavior among college men. [@wardAlcoholUseHostility2025,
@wardAlcoholSexualConsent2012]

#### Perceived-peer differences

Among college men, peer relationships can be critical to their risk of
perpetrating sexual aggression [@swartoutCompanyTheyKeep2013,
@swartoutTrajectoryAnalysisCampus2015] and male peer support for abuse
and coercion have been linked to lower comprehension of sexual consent
and less communicative sexuality. [@wardAlcoholUseHostility2025,
@warrenComprehensionSexualConsent2015] The relevance of a person’s
beliefs about their peers’ ideas of what is normal and acceptable is
best contextualized by the “sociality of consent,” which refers to a
phenomenon wherein sexual consent is not determined privately between
partners but rather extended into the peer group.[@jensen2020]
Qualitative research assessing how young adults use relational and
social contexts to determine consent further underscores the salience of
peer influence, as participants–particularly college women–described
their friends asking them to help interpret and even make judgments
about consent on their behalf after consuming alcohol or other drugs.
[@huntBlurringLineIntoxication2022, @jensen2020,
@marcantonioWeRelyOthers, @marcantonioWhatWouldYou2025]

### Typical and historical behavior

#### Hazardous alcohol use and sexual risk-taking

Alcohol is a salient factor in the context of both sexual assault and
consensual sex among college students, yet sexual assault prevention and
risk-reduction programs rarely include information about how alcohol can
affect consent communication.
[@jozkowskiGenderDifferencesHeterosexual2014] While scholars and
practitioners have started integrating alcohol and sexual assault
programming, [@orchowskiIntegratedAlcoholSexual2023, @gilmore2025] there
is far more research on alcohol-related sexual assault compared to
consent involving alcohol use. Some research suggests that young adults
who engage in binge drinking generally use less active consent
communication than those who do not binge drink,
[@marcantonioEffectsTypicalBinge2022] but the intensity of alcohol
consumption at the encounter level may be more directly relevant to
in-the-moment. Still, cross-sectional surveys cannot establish whether
real-world drinking events decrease the incidence of active consent
cues, and few micro-longitudinal studies have explicitly investigated
situational factors linked to within-person differences in consent
communication.

### Situational correlates of consent cue use

#### Event-level alcohol consumption

As previously mentioned, Willis and colleagues found no significant
event-level associations between alcohol use and consent cues.
[@willisInternalExternalSexual2021] A daily diary dissertation study
found that, among college women engaging in casual sex, Blood Alcohol
Estimate was negatively associated with consent expression, but this
relationship was no longer significant when the sexual behavior type was
included in the model. [@rinioloRoleAlcoholContext2025] One potential
explanation is that an individual’s alcohol consumption is more likely
to result in them communicating consent differently than they typically
would on days when their alcohol use is also atypical–representing a
deviation from their own personal norm. This hypothesis is supported by
previous research which shows that, on days when women drink more than
their own average, they are more likely to experience negative sexual
consequences later that day. [@fordUsingDailyDiary2021, @scaglione2014]

#### Event-level perceived intoxication

Young adults also recognize and describe themselves adjusting the
strategies they use to communicate sexual consent based on their
perceptions of their own intoxication relative to that of a potential
partner, but this recurring theme in qualitative and vignette-based
research has proven difficult to explicate quantitatively.
[@huntBlurringLineIntoxication2022, @laughlinJustOneShot2023] One
hypothesis is that the extent to which perceived differences between
one’s own and their partner's intoxication relate to their use of verbal
consent cues depends on their level of familiarity with that partner.
For example, if someone knows the other person involved in a sexual
encounter well and is motivated to maintain a positive relationship with
them moving forward, they may be more motivated to voice their own
consent verbally and to check in with that other person about their
degree of intoxication and capacity to consent before proceeding with
sexual activity. This hypothesis is supported by the ‘date rape’
literature, which indicates that sexual assaults involving mutual
alcohol consumption are more likely to take place outside the home and
involve someone less well known to the victim.
[@lorenzAlcoholSexualAssault2016]

#### Partner familiarity and sexual precedent

The relationship or degree of familiarity between two people involved in
a sexual encounter has been related to differences in consent
communication, though existing research has produced paradoxical
conclusions regarding the nature of these associations. Sexual
precedence–the idea that engaging in consensual sex at one point in time
implies consent can be assumed for future encounters with that same
person–is supported by research showing people perceive a greater need
to directly communicate about consent when partners are not in a
committed relationship. [@humphreysPerceptionsSexualConsent2007] Yet,
cross-sectional survey research among U.S. college students indicates
that explicit verbal consent cues are less likely to be used when the
sexual relationship is casual. Internal consent feelings (safety,
comfort) tend to be higher for encounters with committed dating partners
and lower for one-time partners, [@walshDualMeasuresSexual2019] and
internal consent feelings are strong predictors of consent cue use, but
patterns observed over time within relationships suggest a complex
relationship between shared sexual history and consent communication
that warrants further evaluation. [@willisSexualPrecedentsEffect2019]

### Significance of the current research

Conflicting findings and gaps in the literature highlight a pressing
need for research grounded in real-time lived experiences that can help
to explain what relational and situational factors are linked to
differences in the degree to which university students’ use different
consent cues. Experiential sampling via daily reports can facilitate a
more nuanced understanding of the types of alcohol involvement that
weaken verbal and nonverbal consent communication during students’
sexual encounters. Knowledge of these contextual factors can then be
used to inform ecological momentary interventions designed to be
delivered when and where the evidence suggests they will be most
effective–an approach that has shown promise for reducing risk of sexual
revictimization among college women.[@maloney2024]

### Research aims and hypotheses

Before addressing primary study aims, our preliminary goal was to
evaluate relationships between sexual event contextual characteristics
and cue use for communicating about consent during sexual encounters.
Gender, for instance, is a significant factor,
[@willisBarriersSuccessAffirmative2018] but there is inconsistent
evidence regarding its associations with consent cues. Some research
suggests men are less likely to use nonverbal cues and more likely to
use verbal cues when compared to women or queer individuals, yet other
studies observe interactions or trends that imply gender differences
hinge on context.
[@marcantonioWithinPersonExaminationAlcoholInvolved2024,
@willisSexualConsentCommitted2021] Previous research also has identified
relational factors with theorized links to alcohol and sexual consent
communication,[@marcantonio2025] but results characterizing the
magnitude and direction of these associations has been mixed. Unpacking
complexity around these person-level and situational factors could yield
valuable information about the conditions under which particular
individuals are more or less likely to use certain types of consent
cues.

**We tested the following preliminary hypotheses to clarify whether
sexual event contextual factors covary with cue use in the current
sample.**

-   H0a: Verbal consent cue use will be greater for encounters involving
    penetrative sexual behaviors when compared to encounters that did
    not advance beyond kissing/touching.

-   H0b: Nonverbal consent cue use will be greater for encounters
    involving more familiar partners when compared to encounters
    involving partners with less familiarity.

-   H0c: Nonverbal consent cue use will be greater for women and gender
    nonconforming (GNC) individuals when compared to men.

**Aim 1: To examine associations between situational alcohol involvement
and consent cue use during sexual encounters.**

-   H1a: Alcohol consumption (# of drinks) will be significantly,
    negatively associated with both verbal and nonverbal consent cue
    use.

-   H1b: The continuous absolute difference between subjective
    participant and perceived partner intoxication ("intoxication
    disparity") will be significantly, negatively associated with verbal
    cue use.

-   H1c: Drinking venue attendance will be significantly, negatively
    associated with both nonverbal and verbal cue use. We expect this
    negative effect will be more pronounced for nonverbal cues.

-   H1d: Among alcohol-involved encounters, there will be an inverted-U
    relationship between both alcohol consumption and subjective
    intoxication and verbal cue use: low → moderate = verbal consent
    increases; moderate → high = verbal consent decreases.

**Aim 2:** quarto publish gh-pages

-   H2a: The negative association between alcohol consumption and verbal
    cue use will be weaker for penetrative behaviors and stronger for
    encounters involving kissing/touching.

-   H2b: The negative association between drinking venue attendance and
    nonverbal cue use will be stronger among less familiar partners.

-   H2c: The negative association between intoxication disparity and
    verbal cue use will be stronger among men compared to women and GNC
    individuals.

**Aim 3: To evaluate the extent to which distal individual factors
predict consent cue use after controlling for partner familiarity and
sexual behavior.**

-   H3a: Social desirability bias and rape myth rejection will be
    positively associated with verbal cues after controlling for partner
    familiarity and behavior type.

-   H3b: Permissive individual and perceived peer attitudes toward
    alcohol and coercion will be negatively associated with verbal cues
    after controlling for partner familiarity and behavior type.

-   H3c: Hazardous alcohol consumption will be negatively associated
    with both verbal and nonverbal cues after controlling for partner
    familiarity and behavior type.

-   H3d: Permissive attitudes toward alcohol and coercion will
    strengthen the negative association between intoxication disparity
    and verbal cue use after controlling for partner familiarity and
    behavior type.

## Method

The current research used baseline and 28-day daily diary data to move
beyond the limitations of prior studies, allowing us to consider
different forms of alcohol involvement in relation to sexual event
context, individual difference factors, and verbal and nonverbal sexual
consent cue use. Any distal, person-level predictors identified as
important will have their relative contributions verified in a model
that also includes proximal situational factors, increasing confidence
that these are reliable targets for intervention or further theory
development. Our lab collected these data from undergraduates at a large
public university in the midwestern U.S. during the Fall 2022 semester.

### Participants

We used Sona Systems to recruit participants from an online subject pool
of undergraduates taking introductory psychology courses. We invited
eligible students to participate via email. Participants were made aware
that the study had two parts: a 30-minute baseline survey, and up to 28
daily diary surveys. At the end of the baseline survey, students were
asked whether they had a smartphone AND whether they wanted to opt into
the daily diary portion of the study. Those who responded affirmatively
to both were enrolled in and pushed daily survey links through SMS for
28 days. 

Eligible participants were those 18 years or older who owned a
smartphone and reported consuming at least one alcoholic beverage in the
past year. A total of 386 students met eligibility criteria and
completed at least 20% of the baseline assessment. Of these, 278
students participated in the daily diary portion of the study.
Participant ages ranged from 18 - 23.

### Procedures

The research team used Qualtrics to construct a series of data
collection instruments for web-administration, and SurveySignal to send
daily text messages with links to each daily diary survey. After
providing informed consent, participants gained access to the baseline
assessment via Qualtrics. After completing the baseline, each
participant received (1.0) class credit as compensation, which
translates to an additional percentage point on their final grade in
their introductory psychology course. Students with a smartphone were
then given the option to participate in the daily diary portion of the
study. Every day for 28 days, each participant who opted in received a
text message between 11:00 AM and 2:00 PM with a link to that day’s
survey. Each day, those who had not yet opened the survey link also
received a reminder text message two hours after the initial text was
sent. For each complete daily diary, participants received 0.25 class
credits as compensation. To incentivize responding, participants who
completed all 28 daily diary surveys received a \$20 Amazon.com gift
card in addition to the 7.0 extra credit percentage points (28 x 0.25
class credits).

The baseline and daily diary surveys were designed to capture
theoretically and empirically supported personal and situational
factors, respectively. The measures we selected were drawn from
peer-reviewed research examining sexual consent and assault.

## Measures

### Baseline

Demographic measures assessed gender, age, race, sexual orientation,
year in school, relationship status, campus involvement (e.g., Greek
life, athletics), residence (e.g., dormitories), disability status, and
socioeconomic status indicators (e.g., Pell grant recipient). Rape myth
rejection was assessed with the Illinois Rape Myths Acceptance scale
[@mcmahonUpdatedMeasureAssessing2011]. Items from the Alcohol and Sexual
Consent scale,[@wardAlcoholSexualConsent2012] the Process-Based Consent
Scale (subtle coercion sub-scale), [@glaceMovingNewModel2020] and the
Rape Attitudes and Beliefs Scale [@burgess2007] were used to assess
individual and perceived peer attitudes toward alcohol and sexual
coercion. Hazardous alcohol use was assessed with the Alcohol Use
Disorders Identification Test [@saundersDevelopmentAlcoholUse1993] and
typical consent behavior was assessed with the Sexual Consent
Scale-Revised. [@humphreysSexualConsentScalerevised2010]

### Daily Diary

Each daily diary prompted participants to report on behaviors and
experiences since their last survey and began with five behavioral
screening items assessing (1) partnered sexual behavior, (2) alcohol use
behavior, (3) online sexual behavior, (4) solitary sexual behavior, and
(5) drug use behavior. We designed the daily surveys so that each
participant would complete 1-2 survey blocks per day assigned based on
their responses to the screening items. When a participant responded
“No” or “Prefer not to respond” to all five screening items with
corresponding behavior blocks, they were assigned to a general mental
health block. Survey blocks were all comparable in length and focused on
sensitive topics to discourage selective responding and ensure that each
person received roughly the same number of follow-up questions per day,
regardless of the experiences they endorsed on the screening items.

Survey blocks were rank-ordered from 1-5 and prioritized partnered
sexual behavior and alcohol use behavior data, such that, whenever a
participant responded “Yes” to more than two screening items, branching
logic directed them to the two-highest ranked survey blocks. The
screening item for partnered sexual behavior read, “Since the last
survey, have you hooked up, engaged in any sexual activity, or tried to
engage in any sexual activity with another person (including flirting,
kissing, touching, or intercourse)?” Response options were: Yes (= 1),
No (= 0), Prefer not to respond (= 3). 

#### Sexual event context

**Partner familiarity:** Participants responded to the question, "How
did you know this other person?" by selecting one of the following
familiarity levels from a list of response options: “I didn’t, we just
met that day;” “They are an acquaintance;” “They are a friend;” and
“They are my current or past dating/intimate partner.” We collapsed the
'acquaintance' and 'friend' response categories into one level, which
was compared against dating/intimate partners and people who just met
via factor variable `fam_lev`.

**Sexual precedent:** Participants responded to the question, “Had you
ever had sexual contact with this person in the past?” with response
options Yes (=1) and No (=0). 

**Sexual behavior type:** Participants were instructed to select all
that apply from a list of specific behaviors that captures escalating
levels of physical intimacy aligned with previously established sexual
hierarchies: Kissing (=1); Touching (=2); Penetration with fingers or
fingering (=3); Oral sex (=4); Vaginal sex (=5); and Anal sex
(=6).[@sanders2010] An additional response option read, “I tried but no
sexual contact happened,” and encounters for which this was the only
response will not be included in the present analysis. A dichotomous
variable comparing events involving Contact (kissing, touching) with
those involving Penetration was created; the Contact v. Penetration
distinction is what differentiates Sexual Assault from Rape in policy
and criminal justice contexts as well as within the scientific
literature. Limited representation across sexual behavior types
precluded the use of an ordinal event-level variable during hypothesis
testing.

#### Alcohol involvement

**Perceived intoxication:** Participants rated their own intoxication
("How intoxicated did you feel right before or during this sexual
activity?") and their perception of their partner’s intoxication ("How
intoxicated do you think the other person felt right before or during
this sexual activity?") using a sliding scale from “Least intoxicated
ever felt” (=0) to “Most most intoxicated ever felt” (=100). 

**Intoxication disparity:** A continuous intoxication "disparity"
variable was calculated as the absolute difference between self-reported
intoxication and perceived partner intoxication. Larger values indicate
lower perceived alignment between self- and partner- intoxication. We
also created a factor variable with categories for equal intoxication
(Parity), for when the participant perceived themselves as more
intoxicated (Self-Higher), and for when the participant perceived the
other person as more intoxicated (Partner-Higher). 

**Pre-encounter alcohol consumption:** Participants were asked, “How
many drinks did you consume prior to or during this activity?” with
response options for the following drink counts: 0; 1-2; 3-4; 5-6; 7-9;
and 10+ drinks. A continuous `drink` variable was created from this
item, ranging from 0-5.

**Drinking venue attendance:** Participants were asked, "Since the last
survey, have you been to a bar, club, or party?" with response options
for "Yes" (=1) and "No" (=0). This item was included in the first survey
block and presented to participants after they completed the five
behavioral screeners, but before branching logic routed them to their
behavior-specific blocks for that daily report.

### Outcomes

**Consent cue use:** After reading the question stem, “To what extent
did you do the following during the sexual activity?” participants rated
the following three items on a unipolar, 11-point sliding scale ranging
from 0 (“Not at all”) to 10 (“Very much”).
[@willisDevelopingValidFeasible2021]

![Mobile view of consent cue items in daily
survey.](images/cues_mobile.png){width="315"}

-   **Verbal:** “Directly asked your partner or communicated verbally
    with your partner about consent.”
-   **Nonverbal, Direct:** “Used direct nonverbal signals to communicate
    with your partner about consent for the sexual activity.”
-   **Nonverbal, Indirect:** “Used subtle signals to communicate with
    your partner about consent for the sexual activity.”

#### Validity of outcome measures

Despite the need for and potential benefits of using experience sampling
methods (ESM) to examine alcohol-involved sexual consent experiences,
there also are unique challenges associated with this approach. One such
challenge is the increased commitment and investment of effort that
repeatedly completing assessments requires from
participants.[@vanhooff2007] ESM researchers can mitigate participant
burden by making surveys brief, a common but imperfect solution that
frequently necessitates measuring variables with one or just a few items
per construct.[@fisher2012] Measures composed of only one or two items
are often criticized by researchers with well-founded concerns that they
might be unreliable, more susceptible to measurement error, or fail to
capture complex constructs. [@robins2001] Still, other researchers argue
that relying on single-item measures is acceptable when these items are
carefully developed and evaluated for multiple sources of measurement
validity.[@vanhooff2007, @wolfers]

Recent recommendations from the field of communication science propose
three central criteria for validating single-item measures in an ESM
context.

The item must:

1.  Fit the core of a construct’s definition to ensure that it reflects
    the construct well.
2.  Be sensitive to situational changes that demonstrate predictable
    within-person variability
3.  Be sensitive to differences between groups at the between-person
    level

In the sections that follow, I present findings aligned with these
criteria that offer support for the validity of the brief outcome
measures used in the current research.

##### 1) Definitional correspondence

Active consent communication is the use of words or behaviors (i.e.,
cues) to communicate to another person agreement to engage in sexual
activity. It is also referred to as “external” consent communication
because these words and behaviors are used to outwardly express one’s
internal feelings of willingness to participate in the sexual behavior.
The only validated external consent measure that existed until recently
was the 18-item External Consent Scale, which was developed for
cross-sectional research to ask about the respondent’s most recent
sexual encounter (which could have been days to weeks in the
past).[@jozkowskiConsentingSexualActivity2014,
@walshDualMeasuresSexual2019] Because including 18 items in ESM studies
would be time-intensive and burdensome, Willis and colleagues used a
two-phase procedure to develop and evaluate brief measures of external
consent cues specifically designed for use in ESM
studies.[@willisDevelopingValidFeasible2021]

In Phase I, the research team sought to develop items that best
represented the constructs of interest while demonstrating both face
validity and content validity. To do so, operational definitions for the
external and internal consent constructs were derived from seminal
theoretical research and systematic reviews of the sexual consent
literature. Cognitive interviews were then conducted with sexually
active adults (n = 10) using concurrent probing to assess sample item
comprehension, preferred wording, and interpretation. Three
content-matter experts from the field of sexual consent research and
three measurement experts with doctoral training in psychometrics also
provided ratings to determine the degree to which sample items captured
their corresponding operational definitions. The research team then
calculated item-objective congruence (IIOC) values for content alignment
(cutoff .75) to establish which prospective items had higher content
validity.[@turner2003] Items were selected for Phase II by balancing
participants’ judgments and experts’ IIIOC
ratings.[@willisDevelopingValidFeasible2021]

The brief measure of external
consent[@willisDevelopingValidFeasible2021] consists of four items that
were developed to measure four types of active consent cues identified
through previous research: verbal (“I verbally communicated…”),
nonverbal (“I nonverbally communicated…”), explicit (i.e., direct; “I
used straightforward signals to communicate…”) and implicit (ie.,
indirect; “I used subtle signals to communicate…”). Researchers using
the measure have tried to operationalize active external consent using a
composite score derived from participants’ responses to these four
items, with variable success. Across three studies by Shi and colleagues
using three separate datasets, their team found that the four-item
composites consistently had poor reliability (αs \<
.60).[@shiImpactSexualConsent2025] This pattern of findings aligned with
that of another study by Willis and colleagues, wherein internal
reliability was not strong for a composite including verbal, nonverbal,
explicit, and implicit consent cues.
[@willisAssessingWithinPersonVariability2021] Moreover, across the
Shi[@shiImpactSexualConsent2025] and
Willis[@willisAssessingWithinPersonVariability2021] studies, verbal cues
were highly positively correlated with explicit cues, and nonverbal cues
were highly positively correlated with implicit cues.

In the pilot study that was initially conducted to validate the external
consent measure, the research team solicited open-ended feedback from
participants. When queried about whether any of the survey questions
sounded awkward or lacked clarity, a few participants expressed
confusion regarding the explicit item’s phrasing (e.g., “It is unclear
to me what this question (on straightforward signals) means”), as well
as uncertainty about the distinction between ‘nonverbal’ and ‘subtle’
communication (e.g., “how is nonverbal different than ‘subtle’
communication?”, “What differentiates the
two?”).[@willisDevelopingValidFeasible2021] Considering pilot
participant feedback, strong positive correlations between verbal and
explicit cues reported in previous studies,[@shiImpactSexualConsent2025,
@willisAssessingWithinPersonVariability2021] and the goals of the
current research, we decided against including the
explicit/“straightforward” cue item in our daily diary. We did include
both the nonverbal and implicit cue items, but we observed the same
pattern described in previous research. Specifically, nonverbal cues
were highly positively correlated with implicit cues (r = .73).
Considering this strong correlation and the preregistered strategy for
addressing it reported in prior work,[@shiImpactSexualConsent2025] the
nonverbal and implicit cue items were averaged to create a composite
Nonverbal score. Finally, the internal consistency for the verbal,
nonverbal, and implicit cue items was \< .6 (α = 0.57), while the
Nonverbal (composite) and Verbal (single-item) scores were only weakly
correlated (r = .12). Given the insufficient empirical support for
combining all three items and the goals of the current research, this
dissertation examined Nonverbal and Verbal cue use separately. Higher
scores indicated greater use of each type of cue to communicate consent.

##### 2) Situational sensitivity and within-person validity

To meet this criterion, event-level scores should change in predictable
ways across situations within the same person.

In Phase II, items derived and refined in Phase I were administered in a
seven-day pilot study, with the ESM survey sent to participants (*n* =
12) four times per day. Following the study period, pilot participants
completed an exit survey that collected feedback on the sexual consent
measures and assessed their acceptability in the broader context of the
ESM study design. Findings included within-person associations between
internal and external sexual consent that were consistent with
preexisting theory and research asserting these are separable but
related facets of a multidimensional process
[@muehlenhardComplexitiesSexualConsent2016]. For instance, verbal and
nonverbal consent cues were both positively associated with internal
consent feelings of safety/comfort and readiness; verbal cue use was
also positively associated with internal feelings of agreement/want,
while nonverbal cue use was not. [@willisDevelopingValidFeasible2021]
Further, external consent communication was not associated with alcohol
use at the encounter level for participants in committed relationships,
replicating discriminant patterns observed in previous research.
[@jozkowskiDoesDrinkingAlcohol2015]

Recent recommendations for validating single-item ESM measures call for
concurrent associations at the within-person level, as well as
situational sensitivity that demonstrates predictable within-person
changes linked to theory-driven contextual correlates.[@wolfers] As a
benchmark, within-person concurrent validity is present when a distinct
but theoretically related variable significantly predicts the focal
outcome with a standardized effect size greater than .09.[@wolfers] For
example, one ESM study using multilevel models found that a composite
internal consent feelings score significantly predicted both verbal and
nonverbal consent cue use and that more intimate encounters predicted
higher consent communication within persons (verbal β = .53, p \< .001;
nonverbal β = .45, p \< .001; situational effect β = .32, p \<
.001).[@willisAssessingWithinPersonVariability2021].

To evaluate within-person concurrent validity in the current sample, we
examined whether event-level sexual feelings (a composite of subjective
arousal and pleasure; psb_feel_wp) predicted same-event verbal and
nonverbal consent cue use. Results indicated that when participants
reported higher-than-usual sexual feelings during a sexual encounter,
they also reported using more verbal (β = 0.17, SE = 0.03, t(348) =
5.77, p \< .001, 95% CI \[0.11, 0.23\]) and nonverbal (β = 0.13, SE =
0.03, t(345) = 5.27, p \< .001, 95% CI \[0.08, 0.18\]) consent cues.
These findings provide evidence of within-person concurrent validity,
suggesting that higher situational sexual arousal and pleasure co-occur
with greater active consent communication.

To test event-level situational sensitivity, we examined whether type of
sexual behavior (Contact vs. Penetration; act_c) predicted verbal and
nonverbal consent cue use. Encounters involving penetration (vs.
contact-only) were associated with greater verbal cue use, β = 0.24, SE
= 0.04, t(488) = 5.78, p \< .001, 95% CI \[0.16, 0.32\], and greater
nonverbal cue use, β = 0.10, SE = 0.04, t(459) = 2.80, p = .005, 95% CI
\[0.03, 0.17\]. These results support situational sensitivity by showing
that participants’ consent communication varies appropriately with the
intimacy of the sexual behavior reported.

##### 3) Between-person validity evidence

To satisfy multi-level validity, evidence should demonstrate that the
measure functions both within- and between-persons. The previous section
offered within-person evidence; here we present between-person evidence
and then integrate across levels.

Recent research by Shi and colleagues offers support for between-person
concurrent validity in a sample of Chinese adults in mixed-gender
relationships.[@shiImpactSexualConsent2025] Specifically,
explicit-verbal consent communication significantly predicted greater
sexual satisfaction during participants' most recent sexual event with
their partner (β = .31, SE = .01, $R^2$ = .10, *p* \< .001), and
implict-nonverbal consent communication significantly predicted greater
sexual satisfaction in general (β = .13, SE = .01, $R^2$ = .08, *p* \<
.001).[@shiImpactSexualConsent2025] These findings indicate that
individuals who report higher levels of external consent communication
also report higher sexual satisfaction, consistent with theoretical
expectations.

To evaluate between-person convergent validity in the current sample,
participants’ baseline reports of verbal (`v_on`: “I always ask verbally
for consent at each step of a sexual encounter”) and nonverbal
(`non_vrb_bsl`: “Typically, I communicate sexual consent to my partner
using nonverbal signals and body language”) consent were correlated with
person-level aggregates of the same behaviors from the diary
(`verbal_means`: mean verbal cue use across events; `mean_nonverb`: mean
nonverbal cue use across events). Aggregating diary data between persons
aligned with the level of analysis for the baseline measures and
respected the hierarchical structure.

**Verbal cues.** Participants who reported greater use of verbal consent
cues at baseline also tended to report higher average verbal cue use
across diary-reported encounters, *r*(152) = .54, *p* \< .001.

**Nonverbal cues.** The corresponding association for nonverbal consent
cues was smaller but significant, *r*(152) = .20, *p* \< .001,
indicating a consistent relationship between baseline and event-level
nonverbal consent communication.

## Analytic Strategy

Analyses were conducted in parallel frameworks. We selected Linear
Mixed-Effects Models (LMEM) for primary analyses and fit these
multilevel models using the `lme4` package in R.[@bates2015]

Cumulative Link Mixed Models (CLMMs) were selected for use a secondary
robustness check. These secondary analyses were conducted in R using the
`ordinal` package.[@christensen2015, @christensen2019] CLMMs estimate
the cumulative probability of being at or above each ordered response
category (i.e., level of the outcome variable) and allow for inclusion
of random effects to account for the nonindependence of repeated
observations. This approach also handles non-normal outcomes
appropriately without assuming equal intervals between
categories.[@christensen2018]

Robustness was evaluated based on consistency in the direction and
significance of effects across modeling frameworks.

### Data structure and ICCs

Each participant received one diary survey per day for 28 consecutive
days. Responses were not statistically independent because participants
contributed multiple event-level observations. Daily diary reports that
captured sexual activity involving physical contact with another person
(hereafter referred to as sexual events) were retained for analysis. The
full study period spanned approximately 80 days, during which different
participants sometimes reported sexual events on the same calendar date;
however, no individual participant contributed more than one event for
any given day. Consequently, Level 1 units were sexual-event days nested
within participants (Level 2). Models included random intercepts for
participant `(1|nid)` to model between-person differences in baseline
cue-use propensity. Random slopes for time-varying predictors were
included when supported by identifiable variance. Preliminary LMEMs were
fit with and without random slopes for within-person predictors to
determine the appropriate random-effects structure. Models were compared
via likelihood-ratio tests (LRTs) and information criteria; those with
relatively smaller values for the Akaike Index Criterion (AIC) and
Bayesian Index Criterion (BIC) were considered to fit the data better.

To assess the extent to which participants’ reports of consent cue use
varied within versus between individuals across sexual encounters,
intraclass correlation coefficients (ICCs) were calculated for LMEMs on
the transformed outcomes and via unconditional random-intercept CLMMs
for ordinal outcomes. ICCs quantified the proportion of total variance
in each outcome that was attributable to stable, between-person
differences (range = 0–1); the remaining proportion reflected
within-person variability across repeated observations. Although ICCs
are conceptually identical across modeling frameworks, they must be
estimated on the latent logistic scale for ordinal outcomes analyzed
with CLMMs rather than on the continuous scale used in LMEMs. This
distinction is important because ordinal outcomes do not have a directly
estimated residual variance; instead, the residual variance is fixed at
$\pi^2 / 3$ on the logistic scale. [@nakagawa2017] The variance of the
random intercept ($\sigma^2_{u0}$) was therefore extracted from each
unconditional CLMM and combined with the fixed residual variance on the
latent logistic scale according to the formula below.

$$ICC = \frac{\sigma^2_{u0}}{\sigma^2_{u0} + \pi^2 / 3}$$

### Variable preparation and centering

Time-varying predictors were decomposed into within-person and
between-person components. Specifically, each time-varying predictor
$Xij$ (event $i$ for person $j$) was person-mean centered to compute the
person-mean deviation $X_{ij}^{c} = X_{ij} - \bar{X}_{.j}$, modeled at
Level 1. Person means $\bar{X}_{.j} = \frac{1}{n_j} \sum_i X_{ij}$ were
then grand-mean centered for inclusion in models as Level-2 predictors.
Level-1 categorical variables also were decomposed.

Models then included both the within-person effect $X_{ij}^{c}$ (does a
participant use more consent cues on days when they are above their own
typical level of $X$?) and the between-person effect $\bar{X}_{.j}$ (do
participants with higher typical levels of $X$ use more consent cues
overall?), which yielded separate coefficients. For CLMMs, these
coefficients represent distinct effects on the cumulative log-odds of
endorsing higher cue-use categories.

Time-invariant, person-level variables also were grand-mean centered to
place the model intercept at a meaningful reference and to reduce
collinearity between intercepts and slopes. Continuous predictors were
scaled per SD for interpretability. Finally, we dummy-coded nominal
predictors as factors with a theoretically-neutral referent and
contrast-coded ordered predictors.

### Diagnostics

The continuous event-level outcome variables representing verbal
(`verb`) and nonverbal (`nvrb`) consent cue use were visually and
statistically evaluated for distributional properties. The research team
that developed the consent cue items elected to measure each variable on
a unipolar 11-point sliding scale in accordance with preexisting
recommendations for selecting response formats for ESM
measures.[@willisDevelopingValidFeasible2021, @fisher2012] We used the
same response format, such that each variable was measured on an
11-point sliding scale (0–10; “Not at all” to “Very much”), reflecting
how much participants used verbal and nonverbal cues to communicate
consent during a sexual encounter.

Histograms and summary statistics were inspected to assess normality,
skewness, and potential ceiling or floor effects. Non-normal
distributions were observed for both verbal and nonverbal cues.
Specifically, the distributions were right-skewed, showing pronounced
piling at the maximum value (10), suggesting limited variability among
participants reporting high levels of cue use.

{{< embed notebooks/fig1.qmd#fig-3 >}}

{{< embed notebooks/fig1.qmd#fig-1 >}}

We applied a Yeo–Johnson power transformation to the bounded 0–10
cue-use outcomes to reduce skew and improve adherence to LMEM
distributional assumptions. The Yeo–Johnson transformation is a
monotonic power transformation that accommodates zeros and selects an
optimal power parameter (λ) based on the observed
distribution.[@yeo2000] We then evaluated model assumptions on the
transformed scale using residual diagnostics. No substantial deviations
from normality were observed.

For CLMMs, continuous outcomes were recoded into ordered categories
based on anchor-informed cut points reflecting meaningful qualitative
differences in consent cue use. These categories were not intended to
reproduce the original 0–10 metric, but rather to align the ordinal
models with the semantic structure of the response scale.

{{< embed notebooks/fig1.qmd#fig-2 >}}

{{< embed notebooks/fig1.qmd#fig-4 >}}

In a CLMM, each predictor is assumed to have a constant effect on the
cumulative log-odds of endorsing higher outcome categories across all
response thresholds. Violations occur when a predictor’s influence
differs for lower versus higher category transitions (e.g., between “Not
at all → A little” versus “Quite a bit → Very much”). We evaluated this
proportional-odds assumption using complementary statistical and visual
diagnostic approaches.

First, we conducted a global `nominal_test()` of the proportional-odds
model. This assumption concerns fixed effects only, so we tested it
using a cumulative link model (CLM) without random effects. When a
predictor was flagged during the global screen, we then fit
threshold-specific binary logistic mixed models (GLMMs) to examine
whether fixed-effect slopes varied substantially across cutpoints (e.g.,
“Not at all → A little,” “A little → Somewhat,” etc.).

Specifically, for average physical intimacy (`act_pm_c`), the estimated
log-odds were 1.65, 2.07, 1.73, and 1.73 across successive thresholds,
all p \< .05. Because these coefficients were consistent in direction
and magnitude across thresholds (differences ≤ 0.4), the
proportional-odds assumption was considered tenable. A visual inspection
provided further support for this conclusion.

### Model specification

Each daily survey was distributed between 11:00 AM and 2:00 PM (with a
reminder two hours later) and asked about experiences *since the
previous survey*, yielding roughly 24-hour retrospective reports.
Although event spacing was irregular, the daily diary design may have
introduced predictable temporal structure (e.g., due to gradual study
fatigue). To assess whether design-related patterns could bias
fixed-effect estimates, we evaluated whether day on study accounted for
systematic variation in verbal cue use. A likelihood-ratio test
comparing models with and without study day as a fixed predictor
indicated that including study day did not significantly improve model
fit, $\chi^2(1)= 3.12$, $p = .08$.

To evaluate whether temporal trends varied across participants, we
compared models with and without a random slope for study day. Although
the average effect of study day on verbal consent cue use was not
significant, likelihood-ratio tests indicated significant between-person
variability in day-to-day slopes. Allowing the effect of study day to
vary across individuals significantly improved model fit,
$\chi^2(2)= 14.32$, $p < .001$. Accordingly, a random slope for study
day was retained in all models predicting Yeo-Johnson transformed verbal
cue use to account for individual differences in temporal trajectories.

To examine predictors of verbal and nonverbal consent cue use while
accounting for both within- and between-person variability, we fit
conditional mixed-effects models estimating cue use from person- and
encounter-level characteristics, as specified in each hypothesis set.

#### Addressing convergence problems

Convergence problems occur when the estimation routine fails to identify
a stable solution for model parameters. We based our approach to
addressing these problems on Brauer and Curtin's (2018) guidelines
regarding sequential corrective steps that researchers can take when
LMEM convergence fails. Their recommendations prioritize nonintrusive
steps that preserve the intended model specification before moving on to
classic or corrective remedies that alter the model
structure.[@brauerLinearMixedeffectsModels2018]

Specifically, we used the full, cleaned dataset and retained all
available cases following multiple imputation and data quality
verification in order to include as many sexual events and participants
as possible. We took steps to minimize potential convergence issues
prior to model estimation by carefully screening the data for
distributional irregularities and violations of model assumptions, then
applying Yeo-Johnson transformations to both skewed continuous outcomes.
In addition, all predictors were centered, with time-varying predictors
person-mean centered to reduce multicollinearity in the random-effects
structure. When convergence problems nevertheless emerged, we evaluated
models for potential misspecification, ensuring that random slopes were
only specified for predictors that exhibited within-person variation.
All predictors and covariates were modeled as fixed effects, and a
participant random intercept was included in all models to account for
nonindependence.

When convergence difficulties persisted, we proceeded through the
following steps in order. Specifically, models were refit using the
*bobyqa* optimization algorithm, and the maximum number of optimizer
iterations was increased (up to 200,000 function evaluations) to allow
adequate exploration of the likelihood surface. We also evaluated the
data for extremely sparse categories and units with very few
observations in key design cells, then addressed these issues by
collapsing or removing problematic units as needed. Finally, we were
prepared to rescale predictors and, if necessary, the outcomes to
improve numerical stability, but all convergence problems were resolved
before this remedy was required.

## Results

### Descriptive Statistics

Eligible participants included in the analytic sample (n = 152)
completed an average of 23.9 daily surveys (median = 26, SD = 6.07,
range = 2–28), corresponding to an average completion rate of 85.3%.
These data yielded 528 partnered sexual events across the diary period.
Approximately 5% of these events (n = 27) were excluded from analysis
because the participant reported trying to engage in sexual activity
with another person (including flirting, kissing, touching, or
intercourse) but no sexual contact occurred. The analytic sample that
remained contained 501 partnered encounters. The data were structured in
long format, with one row per event and multiple events nested within
participants. A majority of participants (60.50%) contributed more than
one event, with a median of two events per participant and up to 19
events for some individuals. Approximately 60.48% (n = 303) of these
sexual events involved oral, vaginal, or anal penetration, whereas
roughly 39.52% (n = 198) involved kissing or touching.

### Unconditional Multilevel Models

We tested unconditional multilevel models to estimate the ICCs for
verbal and nonverbal sexual consent. ICCs for each cue type are
presented in @tbl-icc. For ordinal verbal cue use, the ICC was .575,
indicating that approximately 57.5% of the total variance was due to
between-person differences and 42.5% to within-person fluctuations
across encounters. These results confirm that participants’ verbal
consent communication varied both across individuals and within
individuals across contexts, supporting the use of multilevel modeling
for subsequent analyses.
```{r, echo=FALSE}
knitr::opts_chunk$set(dev = "png")
```

```{r, echo=FALSE}
library(tidyverse)
library(glue)

source(here::here("_scripts/inline_results.R"))
coef_env <- read_coef_tables(here::here("objects"))
```

{{< embed notebooks/tables.qmd#tbl-icc >}}

### Sexual event contextual covariates

**Hypothesis 0a:** Verbal consent cue use was significantly higher
during penetrative sexual encounters compared to encounters that did not
progress beyond kissing or touching, supporting H0a. This effect was
significant within persons--when a given participant engaged in oral,
vaginal, or anal sex, they reported greater use of verbal consent cues
than during their own contact-only encounters
(`r inline_effect(coef_env, "h01", "lmem", "act_penetration_wp")`).
Although the magnitude of this effect varied across individuals, the
average within-person association was positive.

**Hypothesis 0b:** Encounters occurring in the context of shared
romantic history (i.e., past or present dating partner) involved higher
nonverbal consent cue use than encounters occurring in less familiar
relational contexts when the partners just met that same day
(`r inline_effect(coef_env, "h02", "lmem", "fam3_romantic_pm_c")`).

Contrary to **Hypothesis 0c,** levels of nonverbal consent cue use were
not significantly lower for men when compared to women and gender
nonconforming (GNC) individuals
(`r inline_effect(coef_env, "h03", "lmem", "gend_diMen")`).

------------------------------------------------------------------------

### Alcohol and consent cue use

**Hypothesis 1a** predicted that alcohol consumption (indexed by number
of drinks) would be negatively associated with both verbal and nonverbal
consent cue use. To test this hypothesis, LMEMs were fit separately for
verbal and nonverbal consent outcomes, decomposing alcohol consumption
into within-person (event-level) and between-person (person-mean)
components and including random intercepts for participants. LMEM
results partially supported H1a for verbal consent. There was a
significant, negative between-person effect of alcohol consumption on
verbal cue use, such that participants who generally drank more than
others before sexual events reported lower verbal cue use overall
(`r inline_effect(coef_env, "vh1a", "lmem", "drink_pm_c")`). In
contrast, the within-person effect of alcohol consumption on verbal cue
use was not significant
(`r inline_effect(coef_env, "vh1a", "lmem", "drink_wp")`).

To evaluate whether these conclusions were robust to the ordinal
measurement of verbal consent cue use, we also fit a CLMM predicting
ordinal verbal cue use from the same within- and between-person
components of alcohol consumption, with a random intercept for
participant (logit link). Consistent with LMEM results, the
within-person effect of alcohol consumption was not significantly
associated with verbal cue use
(`r inline_effect(coef_env, "vh1a", "clmm", "drink_wp")`). However, also
consistent with LMEM results, the between-person effect was negative and
statistically significant
(`r inline_effect(coef_env, "vh1a", "clmm", "drink_pm_c")`). This
pattern of findings might suggest that participants who tend to drink
more before sexual encounters typically have lower odds of using higher
levels of verbal consent across events.

H1a was not supported for nonverbal consent. LMEM results indicated that
neither within—
(`r inline_effect(coef_env, "vh1a", "lmem", "drink_wp")`) nor between—
(`r inline_effect(coef_env, "Nvh1a", "lmem", "drink_pm_c")`) person
effects of alcohol consumption on nonverbal cues were significant. The
same pattern was observed among CLMM results for both within—
(`r inline_effect(coef_env, "Nvh1a", "clmm", "drink_wp")`) and between—
(`r inline_effect(coef_env, "Nvh1a", "clmm", "drink_pm_c")`) person
effects of alcohol consumption on nonverbal cues.

**Hypothesis 1b** predicted that greater intoxication disparity—defined
as the absolute difference between participants’ subjective intoxication
and their perception of their partner’s intoxication—would be negatively
associated with verbal consent cue use. LMEM results partially supported
this hypothesis. Between-person differences in intoxication disparity
were significantly associated with lower verbal consent cue use
(`r inline_effect(coef_env, "h1b", "lmem", "parity_con_pm_c")`).
However, the within-person effect of intoxication disparity was not
significant
(`r inline_effect(coef_env, "h1b", "lmem", "parity_con_wp")`).

The secondary robustness check via CLMM yielded consistent results:
there was a significant, negative between-person effect of intoxication
disparity on verbal cue use
(`r inline_effect(coef_env, "h1b", "clmm", "parity_con_pm_c")`), but the
within-person effect of intoxication disparity was not significant
(`r inline_effect(coef_env, "h1b", "clmm", "parity_con_wp")`). This
pattern of findings might suggest that individuals who routinely find
themselves in sexual situations with mismatched intoxication levels tend
to use less verbal consent in general.

**Hypothesis 1c** predicted that attendance at bars or parties would be
negatively associated with both verbal and nonverbal consent cue use,
with stronger effects expected for nonverbal cues. LMEM results
partially supported this hypothesis. Between-person differences in
same-day drinking venue attendance were significantly negatively
associated with cue use during sexual encounters, and this was true for
both nonverbal
(`r inline_effect(coef_env, "nvh1c", "lmem", "bpa_true_pm_c")`) and
verbal (`r inline_effect(coef_env, "vh1c", "lmem", "bpa_true_pm_c")`)
cues. The within-person effects of drinking venue attendance were not
significant for nonverbal
(`r inline_effect(coef_env, "nvh1c", "lmem", "bpa_true_wp")`) or verbal
(`r inline_effect(coef_env, "vh1c", "lmem", "bpa_true_wp")`) consent.

Similarly, across CLMMs, between-person differences in bar-or-party
attendance on sexual encounter days were significantly, negatively
linked to both nonverbal
(`r inline_effect(coef_env, "nvh1c", "clmm", "bpa_true_pm_c")`) and
verbal (`r inline_effect(coef_env, "vh1c", "clmm", "bpa_true_pm_c")`)
cue use, but no significant effects were observed for within-person
deviations predicting nonverbal
(`r inline_effect(coef_env, "nvh1c", "clmm", "bpa_true_wp")`) or verbal
(`r inline_effect(coef_env, "vh1c", "clmm", "bpa_true_wp")`) consent.

### Modifiers of alcohol-consent associations

**Hypothesis 2a** predicted that the negative association between
alcohol consumption and verbal consent cue use would be weaker for
penetrative behaviors and stronger for encounters involving kissing or
touching. To test this hypothesis, interaction terms between alcohol
consumption and sexual behavior type were included in mixed-effects
models predicting verbal consent cue use. Contrary to Hypothesis 2a, the
interaction between alcohol consumption and behavior type was not
statistically significant at either the within-
(`r inline_effect(coef_env, "h2a", "lmem", "drink_wp:s_actPenetration")`)
or between-
(`r inline_effect(coef_env, "h2a", "lmem", "s_actPenetration:drink_pm_c")`)
person level, suggesting the association between pre-encounter number of
drinks consumed and verbal consent cue use did not differ by behavior
type in this sample.

The secondary robustness check via CLMM yielded consistent results:
sexual behavior type did not significantly moderate the within-
(`r inline_effect(coef_env, "h2a", "clmm", "drink_wp:s_actPenetration")`)
nor between-
(`r inline_effect(coef_env, "h2a", "clmm", "s_actPenetration:drink_pm_c")`)
person effects of alcohol consumption on ordinal verbal cue use,
suggesting the association between pre-encounter number of drinks
consumed and verbal consent cue use did not differ by degree of physical
intimacy (contact- v. penetrative- sexual behavior type) in this sample.

**Hypothesis 2b** predicted that the negative association between
drinking venue attendance and nonverbal consent cue use would be
stronger among less familiar partners. Mixed-effects models testing
interactions between bar-and-party attendance and partner familiarity
revealed no statistically significant moderation effects at either the
within- or between-person level ($ps$ \> .219). Thus, the association
between drinking venue attendance and nonverbal cue use did not differ
as a function of partner familiarity in this sample.

**Hypothesis 2c** predicted that the negative association between
intoxication disparity and verbal consent cue use would be stronger
among men compared to women and gender-nonconforming individuals.
Mixed-effects models were consistent in that neither the LMEM
(`r inline_effect(coef_env, "h2c", "lmem", "parity_con_wp:gend_diMen")`)
nor the CLMM
(`r inline_effect(coef_env, "h2c", "clmm", "parity_con_wp:gend_diMen")`)
results supported this hypothesis. Interaction terms between
intoxication disparity and gender were not statistically significant,
indicating that the the within-person slope of intoxication disparity
did not differ by gender in this sample.

### Distal individual factors

Consistent with **Hypothesis 3a**, LMEM results revealed that social
desirability bias (`r inline_effect(coef_env, "h3a", "lmem", "bidr_c")`)
and rape myth rejection
(`r inline_effect(coef_env, "h3a", "lmem", "irma_c")`) were
significantly, positively associated with verbal consent cue use after
controlling for partner familiarity and behavior type. Results of the
secondary robustness check also supported H3a, such that CLMM findings
indicated that social desirability bias
(`r inline_effect(coef_env, "h3a", "clmm", "bidr_c")`) and rape myth
rejection \
(`r inline_effect(coef_env, "h3a", "clmm", "irma_c")`) positively
predicted endorsement of verbal cues.

**Hypothesis 3b** predicted that permissive individual and perceived
peer attitudes toward alcohol and coercion will be negatively associated
with verbal cues after controlling for partner familiarity and behavior
type. Results partially supported this hypothesis. Perceived peer
attitudes were significantly, negatively associated with verbal cue use
(`r inline_effect(coef_env, "h3b", "lmem", "ppat_c")`), whereas
permissive *individual* attitudes were not significantly associated with
verbal cues when controlling for all other predictors in the model
(`r inline_effect(coef_env, "h3b", "lmem", "asco_c")`). CLMM results
yielded the same pattern of findings: after controlling for familiarity
and sexual behavior, perceived peer attitudes
(`r inline_effect(coef_env, "h3b", "clmm", "ppat_c")`) significantly
predicted verbal consent cues while personal attitudes
(`r inline_effect(coef_env, "h3b", "clmm", "asco_c")`) did not.

**Hypothesis 3c** predicted that hazardous alcohol consumption would be
negatively associated with both verbal and nonverbal consent cue use
after controlling for partner familiarity and behavior type. H3c was not
supported for either outcome or within either analysis framework. Across
both verbal (`r inline_effect(coef_env, "vh3c", "lmem", "auds_c")`) and
nonverbal (`r inline_effect(coef_env, "nvh3c", "lmem", "auds_c")`) cues,
trait-level hazardous drinking did not uniquely predict consent
communication when the LMEM included contextual characteristics of the
sexual event as covariates. CLMM results yielded the same pattern for
both verbal (`r inline_effect(coef_env, "vh3c", "clmm", "auds_c")`) and
nonverbal (`r inline_effect(coef_env, "nvh3c", "clmm", "auds_c")`) cues.

**Hypothesis 3d** predicted that permissive individual attitudes toward
alcohol and coercion would strengthen the negative association between
intoxication disparity and verbal cue use after controlling for partner
familiarity and behavior type. To test this hypothesis, interaction
terms between intoxication disparity and permissive attitudes were
included in mixed-effects models alongside partner familiarity and
behavior type. Contrary to Hypothesis 3d, the interaction between
disparity and permissive attitudes was not statistically significant at
the within-
(`r inline_effect(coef_env, "h3d", "lmem", "parity_con_wp:asco_c")`) or
between-
(`r inline_effect(coef_env, "h3d", "lmem", "asco_c:parity_con_pm_c")`)
person level, indicating that the negative association between
mismatched intoxication and verbal consent cue use was not modified by
personal attitudes.

CLMM results were consistent with those derived from the primary LMEM.
The within-person slope of intoxication disparity did not differ by
permissive attitudes
(`r inline_effect(coef_env, "h3d", "clmm", "parity_con_wp:asco_c")`),
and the between-person effect of intoxication disparity on verbal cues
did not vary as a function of individual attitudes toward alcohol and
coercion
(`r inline_effect(coef_env, "h3d", "clmm", "asco_c:parity_con_pm_c")`).

## Limitations

The proposed research is not without limitations. First, there were
measures not included in this data collection effort that we intend to
prioritize in the future. Considering this project focuses on active
consent communication, we did not measure passive “no response” signals,
which are commonly reported cues used to infer and communicate consent
among young adults despite risks associated with their ambiguity and
inconsistent associations with internal willingness. Also, we did not
assess refusal cues that participants may have used to indicate lack of
willingness, so it is possible that participants were also reflecting on
behaviors they used to assert boundaries when responding to items
assessing verbal and nonverbal consent cues. Other measures not included
that could have strengthened the current study are perceptions of
partner cue usage, because we cannot speak to how participants inferred
consent from their partners without those data, nor can we know whether
the body language, behaviors, or comments that participants interpreted
as cues were intended to signal consent. In future research, these
limitations can be addressed through dyadic data collection, though
recruiting in pairs introduces a different set of constraints with
respect to feasibility and generalizability. Finally, we did not provide
options to select from or ask participants to report specific examples
of the behaviors that they were classifying as verbal or nonverbal cues,
and we did not provide participants with the operational definitions for
these constructs prior to study initiation. 

Specific characteristics of the current study sample and design also may
limit what can be accomplished with the proposed research. Participants
were recruited via Sona Systems and compensated with extra credit in
their introductory psychology class, resulting in a college-student
sample with 85% of participants in their first year. More than 90% of
the sample fell below age 20, while a sample with more proportional
representation of upperclassmen may have yielded a greater number of
partnered sexual events and drinking episodes for analysis. Further, as
these data were collected from college students at a primarily white
institution in the midwestern U.S., the sample is homogeneous in terms
of race (77% White) and sexual orientation (87% heterosexual).
Directions for future research include more stringent criteria that
require eligible study participants be sexually active; prior sexual
experience was not an inclusion criterion for participating in the
current research, as it is common for young adults’ first partnered
sexual experiences to occur during their first semester at
university.[@follingstad2023] 

With respect to study design, the experience sampling method used to
support this project was a relative strength compared to the primarily
cross-sectional literature in this area. Using electronic diaries
reduced the timeframe for retrospective recall, however, survey timing
and framing still required participants to think backwards when
reporting on behaviors and experiences that preceded sexual events
(e.g., pre-encounter number of drinks) on the previous day. Thus,
results examining event-level variables can only be interpreted as
associational and predictive within the concurrent window. Additionally,
asking participants to rate, for example, a sexual partner’s
intoxication level relative to their own at the same timepoint that we
asked them about sexual consent communication may have increased
susceptibility to social desirability bias, potentially leading
participants to respond in ways they thought would be perceived more
positively by the research team, or in ways that reduced their own
cognitive dissonance around possibly taking advantage of an intoxicated
person or being taken advantage of themselves. In future research, these
limitations could be addressed using ecological momentary assessment
with ≥2 surveys per day or event-contingent sampling, permitting
assessment of contextual variables pre-encounter and concurrent or
post-encounter variables immediately thereafter. Finally, there is risk
of overfitting to sample-specific nuance, as we are fitting and
evaluating our models in the same sample; techniques for estimating
out-of-sample generalizability were considered but ultimately omitted
from the analytic plan, which prioritizes feasibility. Directions for
future research include goals to replicate and extend this research in a
separate college sample of new participants, as well as among adults in
other sociocultural contexts.

## References

::: {#refs}
:::
