---
title: Verbal and nonverbal sexual consent cue use among university students
subtitle: A micro-longitudinal investigation
author:
  - name: Stephanie B. Ward 
    orcid: https://orcid.org/0000-0002-8191-0158
    corresponding: true
    email: sbward@wisc.edu
    affiliations:
      - Department of Psychology, University of Wisconsin-Madison
keywords:
  - Sexual consent
  - Sexual assault
  - College student
  - Alcohol
date: last-modified
bibliography: zotero_lib.bib
csl: american-medical-association.csl
number-sections: false 
editor_options: 
  chunk_output_type: console
editor:
  wrap: 72
---

## SPECIFIC AIMS

Sexual consent is a complex process involving thoughts, feelings, and
behaviors used to determine and express willingness to engage in sexual
activity with a particular person in a particular context; the presence
of consent distinguishes wanted sexual encounters from sexual assault.
[@muehlenhardComplexitiesSexualConsent2016] People can communicate
consent using active consent cues, which may be verbal or nonverbal and
direct or indirect, ranging from a clear and spoken “yes” to subtle body
language. [@willisAssociationsInternalExternal2019] Programs designed to
prevent sexual assault often target consent communication, but a
substantial gap exists between their curriculum and the varied consent
practices that young people use during sexual encounters. [@curtis2017]
Individual factors such as demographics, attitudes,
[@wardAlcoholUseHostility2025] and typical behavior
[@marcantonioEffectsTypicalBinge2022] are linked to differences in
consent cue use. Gender, for instance, is a significant factor,
[@willisBarriersSuccessAffirmative2018] but there is mixed evidence
regarding its associations with consent cues. Some research suggests men
are more likely than women to use verbal cues relative to indirect,
nonverbal cues, while other studies observe interactions or trends that
imply gender differences hinge on
context.[@marcantonioWithinPersonExaminationAlcoholInvolved2024,
@willisSexualConsentCommitted2021] Unpacking the complexity around
person-level and situational factors could illuminate more information
about the conditions under which particular individuals are likely to
use certain types of consent cues.

Research on the contexts in which individuals use certain consent cues
has been limited. This is a problem because situational factors such as
sexual behavior type, co-occurring intoxication, and feelings of arousal
and pleasure are linked to within-person differences in consent cue use.
[@willisInternalExternalSexual2021, @rinioloRoleAlcoholContext2025]
Within-person variability has been shown to account for more than 70% of
the variance in sexual consent cues,
[@willisAssessingWithinPersonVariability2021] indicating that context
may be more informative than individual differences in how people
communicate consent in general. Researchers examining situational
factors related to consent have done so by conducting systematic,
within-person comparisons of two matched events,
[@marcantonioWithinPersonExaminationAlcoholInvolved2024] or by grounding
responses to each person’s most recent sexual encounter, which could
have occurred weeks to months in the past.
[@walshSexualVictimizationEmotion2024] By asking people to remember past
events, these studies risk greater recall bias relative to
micro-longitudinal studies, which assess situational factors closer in
time to when consent communication occurs.

Although diaries and ecological momentary assessment have been used
extensively in other behavioral science research to illuminate important
time-varying situational predictors of health-risk behaviors,
[@serreEcologicalMomentaryAssessment2015, @rodríguez-blanco2018,
@bolger2003] only one published investigation of this kind has been
conducted in the sexual consent research space. Willis and colleagues
surveyed participants three times per day for 28 days and found that
those engaging in more intimate sexual behavior  (i.e., vaginal or anal
penetration) reported more active consent communication than those
engaging in less intimate behaviors, while no significant event-level
associations between alcohol use and verbal or nonverbal consent cue use
were observed. [@willisInternalExternalSexual2021,
@willisAssessingWithinPersonVariability2021] However, this research
examined a limited set of predictors and reflected a specific
subpopulation of adults in committed relationships; the current project
builds on their findings by probing person- and event-level associations
with consent cue use among college students while expanding the range of
possible predictors. Examining these associations in different samples
and in relation to a wider array of variables is a crucial next step
toward isolating more actionable targets for consent communication
interventions. 

To advance the field, this project leverages micro-longitudinal data to
identify proximal situational correlates of verbal and nonverbal consent
cues that explain variance above and beyond the effects of significant
individual factors. I will approximately 500 sexual events reported by
152 college students to extend previous research investigating between-
and within-person variation in consent communication, with the goal of
informing future research and sexual assault prevention efforts focused
on changing consent behavior. Specifically, I propose the following
aims. 

**Aim 1:** Characterize between-person differences in consent cue usage.

We will examine attitudinal and behavioral of consent cue usage and
whether use varies by gender. **Hypotheses:** *(H1a):* More permissive
individual and perceived peer attitudes toward alcohol-involved consent
and sexual coercion/rape will be negatively related to verbal and
nonverbal cue use, particularly among men vs. women. *(H1b):* Sexual
risk behavior, hazardous alcohol use, and typical nonverbal approach to
consent will be negatively related to verbal consent cue use, and these
associations will not differ by gender.

**Aim 2:** Estimate within-person associations between situational factors
and consent cue usage.

We will examine relational, perceptual, and behavioral time-varying
correlates of consent cue usage. **Hypotheses:** *(H2a)* Greater
differences in participant vs. perceived partner intoxication will be
positively related to consent cue use for encounters involving more
familiar partners (current/former dating partner \> friend \>
stranger).  *(H2b)* On days when participants drink more than their own
average, consent cue use will be significantly reduced after accounting
for the effects of aggregate alcohol use across the study period.

**Aim 3:** Test cross-level hypotheses and evaluate whether interactions
improve model fit.

Considering the relative nascency of this literature, we will use
findings from Aims 1 and 2 to determine whether there are interactions
that should be considered in addition to the following
theoretically-driven   **Hypotheses:** *(H3a)* The negative association
between permissive individual attitudes toward alcohol-involved consent
and cue use will be stronger for encounters involving alcohol use.
*(H3b)* Pre-encounter alcohol consumption (-), sexual behavior type (+),
and social desirability bias (+) will be significantly related to verbal
consent, such that lower drink counts, more intimate sexual acts
(involving penetration \> kissing, touching), and higher levels of
impression management/self-deceptive enhancement will positively predict
verbal cue use within the concurrent, 24-hour window.

## BACKGROUND AND SIGNIFICANCE

### Understanding between-person predictors of active consent cues.

Rates of sexual assault on U.S. college campuses remain virtually
unchanged after more than 30 years of dedicated research and
intervention. [@kossScopeRapeIncidence1987,
@kossScopeRapeVictimization2022] Lack of consent is increasingly
recognized as defining characteristic of sexual assault, leading
institutions to incorporate consent-focused education into violence
prevention efforts. [@beresPerspectivesRapepreventionEducators2020,
@beresRethinkingConceptConsent2014] University policy and educational
initiatives promote explicit and enthusiastic consent communication
(“only yes means yes!”), while young people often rely instead on
passive cues (not resisting, not saying no) and assumptions (“We love
each other”) [@willisSexualConsentCommitted2021] to determine consent.
[@marcantonioEffectsTypicalBinge2022,
@jozkowskiGenderDifferencesHeterosexual2014] Sexual consent education
aims to promote the use of active cues by enhancing consent knowledge
and positive attitudes while decreasing rape-supportive attitudes,
presuming that attitudes shape consent behavior. However, these
interventions do not have a clear underlying mechanism of behavior
change, and they rarely assess behavioral outcomes.
[@beresPerspectivesRapepreventionEducators2020,
@burtonTeachingSexualConsent2023] A prerequisite to improving upon the
value of consent education is testing whether these attitudinal
intervention targets predict actual in-the-moment cue use and
contrasting their explanatory power with that of other between-subjects
factors.

### Attitudes and adherence to social norms

#### Individual differences

Rape-supportive attitudes and beliefs have been theorized to explain
variability in how individuals approach consent communication.
[@jozkowskiChapter9Sexual2022] Perhaps the most prominent example in the
sexual assault and consent literatures concerns Rape Myth Acceptance
(RMA), the endorsement of widely-held false beliefs that justify sexual
aggression and place blame on victims.
[@lonswayAttitudinalAntecedentsRape1995, @payneRapeMythAcceptance1999]
RMA is a robust predictor of sexual violence perpetration; it also has
been used as an outcome measure in several sexual consent interventions
delivered to young people in educational settings.
[@burtonTeachingSexualConsent2023, @baker2014] Individuals with high
levels of RMA often hold the belief that victims of sexual assault “led
the other person on” (e.g., with what could be conceived through this
lens as nonverbal consent cues) or provided verbal agreement and “later
regretted it.” These pervasive schemas suggest individuals with high
levels of RMA may pursue sexual contact without first communicating
their willingness or looking for indications of the other person’s
openness to their advances. [@silver2018,
@mcmahonUpdatedMeasureAssessing2011] There also are rape myths that
specifically excuse perpetrator behavior in the context of alcohol
intoxication (e.g., “It shouldn't be considered rape if the person is
drunk and didn't know what they were doing”). Rape myths are distinct
from but related to permissive Alcohol and Sexual Consent (ASC)
attitudes, [@mcmahonUpdatedMeasureAssessing2011] both of which
contradict sexual assault programming messages about, for example, how
communication signals are easily misinterpreted when alcohol is involved
in a sexual situation. Permissive ASC attitudes reflect social norms
that are prevalent on college campuses (e.g., “Consensual drunk sex is a
normal and harmless part of college life”) and reinforce approval of
intoxicated sexual activity without regard for the associated risks. Our
previous research found that permissive ASC attitudes were negatively
related to ongoing and communicative (i.e., “process-based”) consent
behavior among college men. [@wardAlcoholUseHostility2025,
@wardAlcoholSexualConsent2012] *Time permitting, highlight interaction
with perceived peer attitudes to facilitate transition.*

#### Perceived-peer differences

Among college men, peer relationships can be critical to their risk of
perpetrating sexual aggression [@swartoutCompanyTheyKeep2013,
@swartoutTrajectoryAnalysisCampus2015] and male peer support for abuse
and coercion have been linked to lower comprehension of sexual consent
and less communicative sexuality. [@wardAlcoholUseHostility2025,
@warrenComprehensionSexualConsent2015] The relevance of a person’s
beliefs about their peers’ ideas of what is normal and acceptable is
best contextualized by the “sociality of consent,” which refers to a
phenomenon wherein sexual consent is not determined privately between
partners but rather extended into the peer group.[@jensen2020]
Qualitative research assessing how young adults use relational and
social contexts to determine consent further underscores the salience of
peer influence, as participants–particularly college women–described
their friends asking them to help interpret and even make judgments
about consent on their behalf after consuming alcohol or other drugs.
[@huntBlurringLineIntoxication2022, @jensen2020,
@marcantonioWeRelyOthers, @marcantonioWhatWouldYou2025]

### Typical and historical behavior

#### Hazardous alcohol use and sexual risk-taking

Alcohol is a salient factor in the context of both sexual assault and
consensual sex among college students, yet sexual assault prevention and
risk-reduction programs rarely include information about how alcohol can
affect consent communication.
[@jozkowskiGenderDifferencesHeterosexual2014] 21 While scholars and
practitioners have started integrating alcohol and sexual assault
programming, [@orchowskiIntegratedAlcoholSexual2023, @gilmore2025] 27,28
there is far more research on alcohol-related sexual assault compared to
consent involving alcohol use. 22 Some research suggests that young
adults who engage in binge drinking generally use less active consent
communication than those who do not binge drink,
[@marcantonioEffectsTypicalBinge2022] 5 but the intensity of alcohol
consumption at the encounter level may be more directly relevant to
in-the-moment. Still, cross-sectional surveys cannot establish whether
real-world drinking events decrease the incidence of active consent
cues, and few micro-longitudinal studies have explicitly investigated
situational factors linked to within-person differences in consent
communication.

### Situational correlates of within-person variability in cue use

#### Event-level alcohol consumption

As previously mentioned, Willis and colleagues found no significant
event-level associations between alcohol use and consent cues.
[@willisInternalExternalSexual2021] A daily diary dissertation study
found that, among college women engaging in casual sex, Blood Alcohol
Estimate was negatively associated with consent expression, but this
relationship was no longer significant when the sexual behavior type was
included in the model. [@rinioloRoleAlcoholContext2025] One potential
explanation is that an individual’s alcohol consumption is more likely
to result in them communicating consent differently than they typically
would on days when their alcohol use is also atypical–representing a
deviation from their own personal norm. This hypothesis is supported by
previous research which shows that, on days when women drink more than
their own average, they are more likely to experience negative sexual
consequences later that day. [@fordUsingDailyDiary2021, @scaglione2014]

#### Event-level perceived intoxication

Young adults also recognize and describe themselves adjusting the
strategies they use to communicate sexual consent based on their
perceptions of their own intoxication relative to that of a potential
partner, but this recurring theme in qualitative and vignette-based
research has proven difficult to explicate quantitatively.
[@huntBlurringLineIntoxication2022, @laughlinJustOneShot2023] One
hypothesis is that the extent to which perceived differences between
one’s own and their partner's intoxication relate to their use of verbal
consent cues depends on their level of familiarity with that partner.
For example, if someone knows the other person involved in a sexual
encounter well and is motivated to maintain a positive relationship with
them moving forward, they may be more motivated to voice their own
consent verbally and to check in with that other person about their
degree of intoxication and capacity to consent before proceeding with
sexual activity. This hypothesis is supported by the ‘date rape’
literature, which indicates that sexual assaults involving mutual
alcohol consumption are more likely to take place outside the home and
involve someone less well known to the victim.
[@lorenzAlcoholSexualAssault2016]

#### Partner familiarity and sexual precedence

The relationship or degree of familiarity between two people involved in
a sexual encounter has been related to differences in consent
communication, though existing research has produced paradoxical
conclusions regarding the nature of these associations. Sexual
precedence–the idea that engaging in consensual sex at one point in time
implies consent can be assumed for future encounters with that same
person–is supported by research showing people perceive a greater need
to directly communicate about consent when partners are not in a
committed relationship. [@humphreysPerceptionsSexualConsent2007] Yet,
cross-sectional survey research among U.S. college students indicates
that explicit verbal consent cues are less likely to be used when the
sexual relationship is casual. Internal consent feelings (safety,
comfort) tend to be higher for encounters with committed dating partners
and lower for one-time partners, [@walshDualMeasuresSexual2019] and
internal consent feelings are strong predictors of consent cue use, but
patterns observed over time within relationships suggest a complex
relationship between shared sexual history and consent communication
that warrants further evaluation. [@willisSexualPrecedentsEffect2019]

### Significance of proposed research

Conflicting findings and gaps in the literature highlight a pressing
need for research grounded in real-time lived experiences that can help
to explain what relational and situational factors are linked to
differences in the degree to which university students’ use different
consent cues. Experiential sampling via daily reports can facilitate a
more nuanced understanding of the contexts that make verbal and
nonverbal consent communication less likely during students’ sexual
encounters. Knowledge of these contextual factors can then be used to
inform ecological momentary interventions designed to be delivered when
and where the evidence suggests they will be most effective–an approach
that has shown promise for reducing risk of sexual revictimization among
college women.[@maloney2024]

## Method

The proposed research will use baseline and 28-day daily diary data to
move beyond the limitations of prior studies, allowing us to consider
interdisciplinary correlates of active verbal and nonverbal sexual
consent cues across levels of analysis. Any distal, person-level
predictors identified as important will have their relative
contributions verified in a model that also includes proximal
situational factors, increasing confidence that these are reliable
targets for intervention or further theory development. Our lab
collected these data from students at a large public university in the
midwestern U.S. between October and mid-December of 2022.

### Participants

We used Sona Systems to recruit participants from an online subject pool
of undergraduates taking introductory psychology courses. We invited
eligible students to participate via email. Participants were made aware
that the study had two parts: a 30-minute baseline survey, and up to 28
daily diary surveys. At the end of the baseline survey, students were
asked whether they had a smartphone AND whether they wanted to opt into
the daily diary portion of the study. Those who responded affirmatively
to both were enrolled in and pushed daily survey links through SMS for
28 days. 

Eligible participants were those 18 years or older who owned a
smartphone and reported consuming at least one alcoholic beverage in the
past year. A total of 386 students met eligibility criteria and
completed at least 20% of the baseline assessment. Of these, 278
students participated in the daily diary portion of the study.
Participant ages ranged from 18 - 23. Participants who endorsed
partnered sexual activity (n = 152) reported 501 events in total.

### Procedures

The research team used Qualtrics to construct a series of data
collection instruments for web-administration, and SurveySignal to send
daily text messages with links to each daily diary survey. After
providing informed consent, participants gained access to the baseline
assessment via Qualtrics. After completing the baseline, each
participant received (1.0) class credit as compensation, which
translates to an additional percentage point on their final grade in
their introductory psychology course. Students with a smartphone were
then given the option to participate in the daily diary portion of the
study. Every day for 28 days, each participant who opted in received a
text message between 11:00 AM and 2:00 PM with a link to that day’s
survey. Each day, those who had not yet opened the survey link also
received a reminder text message two hours after the initial text was
sent. For each complete daily diary, participants received 0.25 class
credits as compensation. To incentivize responding, participants who
completed all 28 daily diary surveys received a \$20 Amazon.com gift
card in addition to the 7.0 extra credit percentage points (28 x 0.25
class credits).

The baseline and daily diary surveys were designed to capture
theoretically and empirically supported personal and contextual factors,
respectively. The measures we selected were drawn from peer-reviewed
research examining sexual consent and assault.

## Measures

### Baseline

Demographic measures assessed gender, age, race, sexual orientation,
year in school, relationship status, campus involvement (e.g., Greek
life, athletics), residence (e.g., dormitories), disability status, and
socioeconomic indicators (e.g., Pell grant recipient). Individual
attitudes: assessed with the Illinois Rape Myths Acceptance scale (7
items);[@mcmahonUpdatedMeasureAssessing2011] Sociosexual Orientation
Inventory (3 items);[@simpsonIndividualDifferencesSociosexuality1991]
and the subtle coercion subscale of the Process-Based Consent scale.
[@glaceMovingNewModel2020] Perceived peer attitudes: assessed with the
Rape Attitudes and Beliefs Scale (2 items)[@burgess2007] and Alcohol and
Sexual Consent scale (2 items). [@wardAlcoholSexualConsent2012]
Individual behavior: assessed with the Alcohol Use Disorders
Identification Test (10 items);[@saundersDevelopmentAlcoholUse1993]
original and adapted Sexual Risk Behavior Scale (6 items);[@fino2021,
@davisIntersectionMensSexual2018] and Sexual Consent Scale-Revised (3
items). [@humphreysSexualConsentScalerevised2010] Trait emotion
dysregulation was assessed with the Difficulties in Emotion Regulation
Scale (16 items).[@victor2016]

### Daily Diary

Each daily diary prompted participants to report on behaviors and
experiences since their last survey and began with five behavioral
screening items assessing (1) partnered sexual behavior, (2) alcohol use
behavior, (3) online sexual behavior, (4) solitary sexual behavior, and
(5) drug use behavior. We designed the daily surveys so that each
participant would complete 1-2 survey blocks per day assigned based on
their responses to the screening items. When a participant responded
“No” or “Prefer not to respond” to all five screening items with
corresponding behavior blocks, they were assigned to a general mental
health block. Survey blocks were all comparable in length and focused on
sensitive topics to discourage selective responding and ensure that each
person received roughly the same number of follow-up questions per day,
regardless of the experiences they endorsed on the screening items.

Survey blocks were rank-ordered from 1-5 and prioritized partnered
sexual behavior and alcohol use behavior data, such that, whenever a
participant responded “Yes” to more than two screening items, branching
logic directed them to the two-highest ranked survey blocks. The
screening item for partnered sexual behavior read, “Since the last
survey, have you hooked up, engaged in any sexual activity, or tried to
engage in any sexual activity with another person (including flirting,
kissing, touching, or intercourse)?” Response options were: Yes (= 1),
No (= 0), Prefer not to respond (= 3). 

#### Relational

**Partner familiarity:** Participants responded to the question,
"How did you know this other person?" by selecting one of the following
familiarity levels from a list of response options: “I didn’t, we just
met that day;” “They are an acquaintance;” “They are a friend;” and
“They are my current or past dating/intimate partner.”

**Sexual precedence:** Participants responded to the question, “Had you
ever had sexual contact with this person in the past?” with response
options Yes (=1) and No (=0). 

#### Perceptual

**Perceived intoxication:** Participants rated their own intoxication
("How intoxicated did you feel right before or during this sexual
activity?") and their perception of their partner’s intoxication ("How
intoxicated do you think the other person felt right before or during
this sexual activity?") using a sliding scale from “Least intoxicated
ever felt” (=0) to “Most most intoxicated ever felt” (=100). 

**Intoxication disparity:** A continuous intoxication "disparity"
variable will be calculated as the absolute difference between
self-reported intoxication and perceived partner intoxication. Larger
values indicate lower perceived alignment between self- and partner-
intoxication. We also will create a factor variable with categories for
equal intoxication (Parity), for when the participant perceived
themselves as more intoxicated (Self-Higher), and for when the
participant perceived the other person as more intoxicated
(Partner-Higher). 

#### Behavioral

**Sexual behavior type:** Participants were instructed to select all
that apply from a list of specific behaviors that captures escalating
levels of physical intimacy aligned with previously established sexual
hierarchies: Kissing (=1); Touching (=2); Penetration with fingers or
fingering (=3); Oral sex (=4); Vaginal sex (=5); and Anal sex
(=6).[@sanders2010] An additional response option read, “I tried but no
sexual contact happened,” and encounters for which this was the only
response will not be included in the present analysis. A dichotomous
variable comparing events involving Contact (kissing, touching) with
those involving Penetration will be created, as will an ordinal
event-level variable if representation across behaviors permits this
level of granularity. The Contact v. Penetration distinction is what
differentiates Sexual Assault from Rape in policy and criminal justice
contexts as well as within the scientific literature.

**Pre-encounter alcohol consumption:** Participants were asked, “How
many drinks did you consume prior to or during this activity?” with
response options for the following drink counts: 0; 1-2; 3-4; 5-6; 7-9;
and 10+ drinks.

### Outcome variables

**Consent cue use:** After reading the question stem, “To what extent
did you do the following during the sexual activity?” participants rated
the following three items on a unipolar, 11-point sliding scale ranging
from 0 (“Not at all”) to 10 (“Very much”). [@willisDevelopingValidFeasible2021]

![Mobile view of consent cue items in daily
survey.](images/cues_mobile.png){width="315"}

-   **Verbal:** “Directly asked your partner or communicated verbally
    with your partner about consent.”
-   **Nonverbal, Direct:** “Used direct nonverbal signals to communicate
    with your partner about consent for the sexual activity.”
-   **Nonverbal, Indirect:** “Used subtle signals to communicate with
    your partner about consent for the sexual activity.”

### Validity of outcome measures

Despite the need for and potential benefits of using experience sampling
methods (ESM) to examine alcohol-involved sexual consent experiences,
[@marcantonioHowDoes2025] there also are unique challenges associated
with this approach. One such challenge is the increased commitment and
investment of effort that repeatedly completing assessments requires
from participants.[@vanhooff2007] ESM researchers can mitigate
participant burden by making surveys brief, a common but imperfect
solution that frequently necessitates measuring variables with just a
few items per construct.[@fisher2012] Measures composed of only one or
two items are often criticized by researchers with well-founded concerns
that they might be unreliable, more susceptible to measurement error, or
fail to capture complex constructs. [@robins2001] Still, other
researchers demonstrate that repeatedly administering assessments yields
multiple indicators, decreasing random measurement
error,[@shiffmanEcologicalMomentaryAssessment2008,
@csikszentmihalyi2014] and argue that relying on single-item measures is
acceptable when these items are carefully developed and evaluated for
multiple sources of measurement validity.[@vanhooff2007, @wolfers]

In the sections that follow, I review literature detailing how the items
used to measure verbal and nonverbal consent cue use for the proposed
research were originally developed and tested by a separate research
team. I then discuss evidence of measurement validity from previous
research before describing the use of these items and their psychometric
properties in the current sample.

#### Initial development & evaluation

Active consent communication is the use of words or behaviors (i.e.,
cues) to communicate to another person agreement to engage in sexual
activity. It is also referred to as "external" consent communication
because these words and behaviors are used to outwardly express one's
internal feelings of willingness to participate in the sexual behavior.
Although prior research reviewed in this proposal suggests that sexual
consent varies from one encounter to the next, the only validated
consent measures that existed until recently were those developed for
cross-sectional research [@jozkowskiConsentingSexualActivity2014,
@walshDualMeasuresSexual2019]. As such, Willis and colleagues used a
two-phase procedure to develop and evaluate brief measures of internal
consent feelings and external consent cues specifically designed for use
in ESM studies. [@willisDevelopingValidFeasible2021]

In Phase I, the research team sought to develop items that best
represented the constructs of interest while demonstrating both face
validity and content validity. To do so, operational definitions for the
external and internal consent were derived from seminal theoretical
research and systematic reviews of the sexual consent literature.
Cognitive interviews were then conducted with sexually active adults
(*n* = 10) using concurrent probing to assess sample item comprehension,
preferred wording, and interpretation. Three content-matter experts from
the field of sexual consent research and three measurement experts with
doctoral training in psychometrics also provided ratings to determine
the degree to which sample items captured their corresponding
operational definitions. The research team then calculated IIOC values
for content alignment (cutoff .75) to establish which prospective items
had higher content validity. Items were selected for Phase II by
balancing participants' judgments and experts' IIIOC
ratings.[@willisDevelopingValidFeasible2021]

In Phase II, items derived and refined in Phase I were administered in a
seven-day pilot study, with the ESM survey sent to participants (*n* =
12) four times per day. Following the study period, pilot participants
completed an exit survey that collected feedback on the sexual consent
measures and assessed their acceptability in the broader context of the
ESM study design. Findings included within-person associations between
internal and external sexual consent that were consistent with
preexisting theory and research asserting these are separable but
related facets of a multidimensional process 
[@muehlenhardComplexitiesSexualConsent2016]. For instance,
verbal and nonverbal consent cues were both positively associated with
internal consent feelings of safety/comfort and readiness; verbal cue
use was also positively associated with internal feelings of
agreement/want, while nonverbal cue use was not.
[@willisDevelopingValidFeasible2021] Further, external
consent communication was not associated with alcohol use at the
encounter level for participants in committed relationships, replicating
discriminant patterns observed in previous research.
[@jozkowskiDoesDrinkingAlcohol2015] We contextualize these findings and
the pilot participants' feedback in the sections that follow.

#### Preexisting validity evidence

Willis and colleagues established feasibility for ESM administration and
solid content and response-process foundations for their external
consent cue items. [@willisDevelopingValidFeasible2021] However, gaps
remained in the evidence base for these measures' validity following
their initial publication, some of which have since been addressed.

Recent recommendations for validating single-item ESM measures call for
concurrent associations at the within-person level, as well as
situational sensitivity that demonstrates predictable within-person
changes linked to theory-driven contextual correlates.[@wolfers] These
guidelines from the field of communication science suggest evidence
concerning within-person concurrent validity is present when a distinct
but theoretically-related variable significantly predicts the focal
outcome with a standardized effect size greater than .09.[@wolfers] This
was the case in an ESM study wherein the researchers fit multilevel
models and found that a composite internal consent feelings score
significantly predicted verbal (β = .53, *p* \< .001) and nonverbal (β =
.45, *p* \< .001) consent cue use
[@willisAssessingWithinPersonVariability2021]. The same study
demonstrated event-level situational sensitivity to type of sexual
behavior within the same person. Specifically, encounters involving more
intimate sexual behaviors predicted higher consent communication scores
compared to encounters reported by the same person involving less
intimate behaviors (β = .32, *p* \< .001).

To establish multi-level validity of single-item ESM measures, evidence
is required both within and between-persons.[@wolfers] Recent research
by Shi and colleagues offers support for between-person concurrent
validity in a sample of Chinese adults in mixed-gender
relationships.[@shiImpactSexualConsent2025] Specifically,
explicit-verbal consent communication significantly predicted greater
sexual satisfaction during participants' most recent sexual event with
their partner (β = .31, SE = .01, $R^2$ = .10, *p* \< .001), and
implict-nonverbal consent communication significantly predicted greater
sexual satisfaction in general (β = .13, SE = .01, $R^2$ = .08, *p* \<
.001).[@shiImpactSexualConsent2025]

The brief measure of external
consent[@willisDevelopingValidFeasible2021] consists of four items that
were developed to measure four types of active consent cues identified
through previous research: verbal (“I verbally communicated..."),
nonverbal ("I nonverbally communicated..."), explicit (i.e., direct; "I
used straightforward signals to communicate...") and implicit (ie.,
indirect; "I used subtle signals to communicate..."). Researchers using
the measure have tried to operationalize active external consent using a
composite score derived from participants’ responses to these four
items, with variable success. Across three studies by Shi and colleagues
using three separate datasets, their team found that the four-item
composites consistently had poor reliability (αs \<
.60).[@shiImpactSexualConsent2025] This pattern of findings aligned with
that of another study by Willis and colleagues, wherein internal
reliability was not strong for a composite including verbal, nonverbal,
explicit, and implicit consent cues.
[@willisAssessingWithinPersonVariability2021] Moreover, across the
Shi\[@shiImpactSexualConsent2025\] and
Willis\[@willisAssessingWithinPersonVariability2021\] studies, verbal
cues were highly positively correlated with explicit cues, and nonverbal
cues were highly positively correlated with implicit cues.

#### Consent cue items in the current sample

In the pilot study that was initially conducted to validate the internal
and external consent measures, the research team solicited open-ended
feedback from participants. When queried about whether any of the survey
questions sounded awkward or lacked clarity, a few participants
expressed confusion regarding the explicit item's phrasing (e.g., "It is
unclear to me what this question (on straightforward signals) means"),
as well as uncertainty about the distinction between ‘nonverbal’ and
‘subtle’ communication (e.g., "how is nonverbal different than ‘subtle’
communication?", "What differentiates the
two?").[@willisDevelopingValidFeasible2021]

Considering pilot participant feedback, strong positive correlations
between verbal and explicit cues reported in previous
studies,[@shiImpactSexualConsent2025,
@willisAssessingWithinPersonVariability2021] and the goals of the
current research, we decided against including the
explicit/"straightforward" cue item in our daily diary. We did include
both the nonverbal and implicit cue items, but we observed the same
pattern described in previous research. Specifically, nonverbal cues
were highly positively correlated with implicit cues $(r = .73)$.
Considering this strong correlation and the preregistered strategy for
addressing it reported in prior work,[@shiImpactSexualConsent2025] the
nonverbal and implicit cue items were averaged to create a composite
Nonverbal score.

Finally, the internal consistency for the verbal, nonverbal, and
implicit cue items was \< .6 (α = 0.57), while the Nonverbal (composite)
and Verbal (single-item) scores were only weakly correlated $(r = .12)$.
Given the insufficient empirical support for combining all three items
and the goals of the current research, this dissertation will examine
Nonverbal and Verbal cue use separately. Higher scores indicate greater
use of each type of cue to communicate consent.

#### Validity in the current sample

To evaluate between-person convergent validity in the current sample, I
examined the associations between participants’ baseline reports of
verbal (`verbb_ong`; "I always ask verbally for consent at each step of
a sexual encounter") and nonverbal (`non_vrb_bsl`; "Typically, I
communicate sexual consent to my partner using nonverbal signals and
body language") consent, and their average levels of verbal and
nonverbal cue use, respectively, across all sexual encounters reported
in the daily diaries. Higher correlations between corresponding baseline
and diary measures would indicate that individuals who report more
frequent use of a particular type of consent cue in general also tend to
report greater use of that cue across diary-reported events. I
aggregated the event-level diary data to the person level by computing
each participant’s mean verbal (`verbal_means`) and nonverbal
(`mean_nonverb`) cue use, thereby aligning the level of analysis
(between-person) with the baseline measure and respecting the
hierarchical structure of the diary data.

Between-person convergent validity was supported for both verbal and
nonverbal consent cue use. Participants who reported greater use of
verbal consent cues at baseline also tended to report higher average
verbal cue use across diary-reported encounters,
$r(152) = .54, p < .001$. The corresponding association for nonverbal
consent cues was smaller but significant, $r(152) = .20, p < .001$,
indicating a consistent—albeit weaker—relationship between baseline and
event-level nonverbal consent communication.

To evaluate within-person concurrent validity, we examined whether
event-level sexual feelings (a composite of subjective arousal and
pleasure; `psb_feel_wp`) predicted same-event verbal and nonverbal
consent cue use. Results indicated that when participants reported
higher-than-usual sexual feelings during a sexual encounter, they also
reported using more verbal, β = 0.17, SE = 0.03, t(348) = 5.77,
*p*\<.001, 95% CI \[0.11, 0.23\], and nonverbal, β = 0.13, SE = 0.03,
t(345) = 5.27, *p*\<.001, 95% CI \[0.08, 0.18\], consent cues. These
findings provide evidence of within-person concurrent validity,
suggesting that higher situational sexual arousal and pleasure co-occur
with greater active consent communication.

To test event-level situational sensitivity, we examined whether type of
sexual behavior (Contact vs. Penetration; act_c) predicted verbal and
nonverbal consent cue use. Encounters involving penetration (vs.
contact-only) were associated with greater verbal cue use, β = 0.24, SE
= 0.04, t(488) = 5.78, *p*\<.001, 95% CI \[0.16, 0.32\], and greater
nonverbal cue use, β = 0.10, SE = 0.04, t(459) = 2.80, *p* = .005, 95%
CI \[0.03, 0.17\]. These results support situational sensitivity by
showing that participants’ consent communication varies appropriately
with the intimacy of the sexual behavior reported.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Detect and manage conflicts
library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# Load packages
library(tidyverse)
library(readr)
library(dplyr)
library(ordinal)
library(tidyr)
library(ggplot2)
library(lubridate)
library(janitor)
library(lme4)
# library(lmerTest) 
library(broom)
library(parameters)
library(effectsize)

# Set path to data
path_ready <- "_data/data_ready"

# Read in data
d <- read_csv(here::here(path_ready, "ready_merg.csv"), show_col_types = FALSE)

# Listwise deletion for preliminary analyses for proposal defense
d <- d %>% filter(!is.na(psb6))

# Load functions
source(here::here("_scripts/functions.R"))
```

#### Outcome distributions

The continuous event-level outcome variables representing verbal
(ver_pc) and nonverbal (non_verb) consent cue use were visually and
statistically evaluated for distributional properties. Each variable was
measured on a 0–10 scale, reflecting how much participants used verbal
and nonverbal cues to communicate consent during a given sexual
encounter.

Histograms, density plots, and summary statistics were inspected to
assess normality, skewness, and potential ceiling or floor effects. Both
ver_pc and non_verb displayed non-normal distributions.

Specifically, bimodality was observed, with many responses clustering
near the lowest (0) and at upper-most anchor (10).

{{< embed notebooks/fig1.qmd#fig-1 >}}

The distributions were right-skewed, showing pronounced piling at the
maximum value (10), suggesting limited variability among participants
reporting high levels of cue use.

{{< embed notebooks/fig1.qmd#fig-3 >}}

Because the response scales were both bounded (0–10) and anchor-labeled
(“Not at all” to “Very much”), the continuous treatment of these data
risked overstating precision and violating linear model assumptions. To
accommodate the non-normal distribution, both outcomes were recoded into
five ordered categories (i.e., Not at all, A little, Somewhat, Quite a
bit, Very much).

```{r}
d <- d %>%
  mutate(
    verb_bins = cut(
      ver_pc,
      breaks = c(-Inf, 0, 3, 6, 9, Inf),
      labels = c("Not at all", "A little", 
                 "Somewhat", "Quite a bit", "Very much"),
      right = TRUE,
      include.lowest = TRUE),
    verb_bins = ordered(verb_bins,
                        levels = c("Not at all", "A little", 
                                   "Somewhat", "Quite a bit", "Very much")))

# Verify
is.ordered(d$verb_bins)
table(d$verb_bins, useNA = "ifany")
```

```{r}
# Create ordered bins for nonverbal consent cues
d <- d %>%
  mutate(
    nonvb_bins = cut(
      non_verb,
      breaks = c(-Inf, 0, 3, 6, 9, Inf),
      labels = c("Not at all", "A little", 
                 "Somewhat", "Quite a bit", "Very much"),
      right = TRUE,
      include.lowest = TRUE
    ),
    nonvb_bins = ordered(nonvb_bins,
                         levels = c("Not at all", "A little", 
                                    "Somewhat", "Quite a bit", "Very much")))
# Verify
is.ordered(d$nonvb_bins)
table(d$nonvb_bins, useNA = "ifany")
```

This recoding produced ordinal variables `verb_bins` and `nonvb_bins`,
reflecting meaningful, empirically-informed gradations of consent cue
use consistent with the scale anchors.

{{< embed notebooks/fig1.qmd#fig-2 >}}


{{< embed notebooks/fig1.qmd#fig-4 >}}


## Analytic Strategy

Given the ordinal structure and non-normality of the outcome variables,
Cumulative Link Mixed Models (CLMMs) were selected for primary analyses.
CLMMs model the cumulative probabilities of being at or above each
ordered response category and allow us to include random effects to
account for the nonindependence of repeated diary observations within
persons. This approach handles non-normal outcomes appropriately without
assuming equal intervals between categories, facilitating more accurate
and interpretable estimation of predictors' effects on the ordered
likelihood of greater consent cue use.

To demonstrate feasibility and researcher capability, I present
step-by-step examples at each stage of the analytic plan for estimating
models predicting consent cue use. This example uses verbal cue use as
the outcome, though the same plan will be systematically applied across
analyses, with one exception: preliminary modeling for the dissertation
proposal defense will use listwise deletion. In contrast, primary models
will employ multiple imputation to address missing predictor data while
retaining all available outcome information. Ordinal outcome variables
will not be imputed.

### Data structure and ICCs

Each participant received one diary survey per day for 28 consecutive
days. Responses are not statistically independent because participants
contributed multiple event-level observations. Diaries capturing sexual
activity involving physical contact with another person (hereafter
referred to as sexual events) were retained for analysis (≈499). The
full study period spanned approximately 80 days, during which different
participants sometimes reported sexual events on the same calendar date;
however, no individual participant contributed more than one event for
any given day. Consequently, Level 1 units are sexual-event days nested
within participants (Level 2). All CLMMs will include random intercepts
for participant `(1|nid)` to model between-person differences in
baseline cue-use propensity. Random slopes for time-varying predictors
will be included when supported by identifiable variance; models will be
compared via likelihood-ratio tests and AIC.

To assess the extent to which participants’ reports of consent cue use
varied within versus between individuals across sexual encounters,
intraclass correlation coefficients (ICCs) will be computed using
unconditional CLMMs. ICCs quantify the proportion of total variance in
an outcome that is attributable to stable, between-person differences
(range = 0–1), with the remaining proportion reflecting within-person
variability across repeated observations. Although ICCs are conceptually
identical across modeling frameworks, they must be estimated on the
latent logistic scale for ordinal outcomes analyzed with CLMMs rather
than on the continuous scale used in linear mixed-effects models (LMMs).
This distinction is important because ordinal outcomes do not have a
directly estimated residual variance; instead, the residual variance is
fixed at $\pi^2 / 3$ on the logistic scale. [@nakagawa2017]

Unconditional CLMMs with random intercepts for participants (`nid`) and
no predictors will be fit for both types of consent cue use.

```{r}
clmm_null <- clmm(verb_bins ~ 1 + (1 | nid), data = d, # null CLMM
                  link = "logit", Hess = TRUE)
```

The variance of the random intercept ($\sigma^2_{u0}$) will be extracted
from each model and combined with the fixed residual variance on the
latent logistic scale according to the formula presented below.

$$ICC = \frac{\sigma^2_{u0}}{\sigma^2_{u0} + \pi^2 / 3}$$

```{r}
# --- robust variance extraction ---
vc <- VarCorr(clmm_null)
# take the variance matrix for the 'nid' grouping factor
var_u0 <- as.numeric(vc$nid)
# combine with the fixed residual variance on the latent logistic scale
icc_logit <- var_u0 / (var_u0 + (pi^2 / 3))

within_prop <- 1 - icc_logit
sprintf("ICC (between-person): %.3f; Within-person share: %.3f", icc_logit, within_prop)
```

For verbal cue use, the ICC was 0.575, indicating that approximately
57.5% of the total variance was due to between-person differences and
42.5% to within-person fluctuations across encounters. These results
confirm that participants’ verbal consent communication varied both
across individuals and within individuals across contexts, supporting
the use of multilevel modeling for subsequent analyses.

### Variable preparation and centering

Time-varying predictors will be decomposed into within-person and
between-person components. Specifically, each time-varying predictor
$Xij$ (event $i$ for person $j$) will be person-mean centered to compute
the person-mean deviation $X_{ij}^{c} = X_{ij} - \bar{X}_{.j}$, modeled
at Level 1.

```{r}
d <- d %>%
  group_by(nid) %>%
  mutate(
    # person mean (Level-2 component)
    drinks_pm = mean(drinks, na.rm = TRUE),

    # within-person deviation (Level-1 component)
    drinks_wp = drinks - drinks_pm
  ) %>%
  ungroup()

d <- d %>% mutate(psb_act_num = if_else(psb_act == "Penetration", 1, 0))
d <- d %>%
  group_by(nid) %>%
  mutate(
    # Between-person component (proportion of events that involved penetration)
    act_pm = mean(psb_act_num, na.rm = TRUE),

    # Within-person deviation (event value minus person mean)
    act_wp = psb_act_num - act_pm
  ) %>%
  ungroup()

d %>% select(drinks_pm, act_pm, drinks_wp, act_wp) %>% head()
```

Person means $\bar{X}_{.j} = \frac{1}{n_j} \sum_i X_{ij}$ will then be
grand-mean centered for inclusion in models as Level-2 predictors.

```{r}
d <- d %>%
  mutate(
    drinks_pm_c = drinks_pm - mean(drinks_pm, na.rm = TRUE),
    act_pm_c    = act_pm - mean(act_pm, na.rm = TRUE))
```

Level-1 categorical variables also must be decomposed.

```{r}
d <- decompose_factor(d, id = "nid", var = "parity_cat", 
                      baseline = "Equal", prefix = "parity")

d %>%
  group_by(nid) %>%
  summarise(across(ends_with("_wp"), \(x) mean(x, na.rm = TRUE))) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)))
```

The model then includes both the within-person effect $X_{ij}^{c}$ (does
a participant use more consent cues on days when they are above their
own typical level of $X$?) and the between-person effect $\bar{X}_{.j}$
(do participants with higher typical levels of $X$ use more consent cues
overall?), yielding separate coefficients (effects on the cumulative
log-odds of higher consent-cue categories).

```{r}
# check work to confirm person-mean centering was successful
tol <- 1e-10

per_person_means <- d %>%
  group_by(nid) %>%
  summarise(across(ends_with("_wp"), ~ mean(.x, na.rm = TRUE)), .groups = "drop")

# any violations?
violations <- per_person_means %>%
  pivot_longer(-nid, names_to = "var", values_to = "mean_wp") %>%
  filter(!is.na(mean_wp), abs(mean_wp) > tol)

nrow(violations)
```

Time-invariant, person-level variables (e.g., social-desirability bias)
will also be grand-mean centered to place the model intercept at a
meaningful reference (average participant at their average levels) and
to reduce collinearity between intercepts and slopes. Continuous
predictors may be scaled (e.g., per SD) for interpretability.

```{r}
d$bidr_sc <- d$bidr_s - mean(d$bidr_s, na.rm = T) # social desirability bias 
```

Nominal predictors will be dummy-coded with a theoretically neutral
referent; ordered predictors will be treated as ordinal or
contrast-coded depending on theoretical expectations about spacing.

### Diagnostics: Proportional-odds assumption

Analyses will be conducted in R using `ordinal::clmm()`, which
implements maximum-likelihood estimation. Each model’s proportional-odds
assumption will be evaluated through statistical and visual diagnostics.

In a CLMM, each predictor is assumed to have a constant effect on the
cumulative log-odds of endorsing higher outcome categories across all
response thresholds. Violations occur when a predictor’s influence
differs for lower versus higher category transitions (e.g., between “Not
at all → A little” versus “Quite a bit → Very much”). We evaluate this
assumption using complementary approaches.

To evaluate the proportional-odds assumption, I will first conduct a
global `nominal_test()` of the proportional-odds model. Of note, this
assumption is about the fixed effects, so I will test it using a
cumulative link model (CLM) without random effects.

```{r}
# 1) list variables needed for model
vars_pre <- c(
  "verb_bins", # outcome
  "nid", # participant ID
  "drinks_wp", "drinks_pm_c", # drink count, within- and between-
  "act_wp","act_pm_c", # sexual behavior type, within- and between-
  "bidr_sc" # time-invariant, person-level social desirability bias
)

# 2) build a clean analysis dataset with no missing values on any of these
d_clm <- d %>%
  select(all_of(vars_pre)) %>%
  tidyr::drop_na()

# 3) fit the proportional-odds CLM on the complete-case data
prop_tst <- clm(
  verb_bins ~ drinks_pm_c + drinks_wp + act_pm_c + act_wp +
    bidr_sc, data = d_clm, link = "logit")

# 4) run the proportional-odds diagnostic
nominal_test(prop_tst)
```

For predictors flagged during the global screen, such as the between-person effect of sexual behavior type (i.e., high v. low level of physical intimacy), we will then fit
threshold-specific binary logistic mixed models (GLMMs) to examine
whether fixed-effect slopes vary substantially across cutpoints (e.g.,
“Not at all → A little,” “A little → Somewhat,” etc.).

```{r}
d_thr <- d %>%
  mutate(bin_num = as.numeric(verb_bins),
         cum1 = as.integer(bin_num >= 2),
         cum2 = as.integer(bin_num >= 3),
         cum3 = as.integer(bin_num >= 4),
         cum4 = as.integer(bin_num >= 5))

# compare slopes for the between-person effect of sexual behavior type
m1 <- glmer(cum1 ~ act_pm_c + (1|nid), data = d_thr, family = binomial)
m2 <- glmer(cum2 ~ act_pm_c + (1|nid), data = d_thr, family = binomial)
m3 <- glmer(cum3 ~ act_pm_c + (1|nid), data = d_thr, family = binomial)
m4 <- glmer(cum4 ~ act_pm_c + (1|nid), data = d_thr, family = binomial)

tbl <- bind_rows(
  fixef_tbl(m1, "≥ A little"),
  fixef_tbl(m2, "≥ Somewhat"),
  fixef_tbl(m3, "≥ Quite a bit"),
  fixef_tbl(m4, "Very much vs below")
) %>%
  filter(term == "act_pm_c")

tbl
```

For average physical intimacy (`act_pm_c`), the estimated log-odds were
1.65, 2.07, 1.73, and 1.73 across successive thresholds, all p \< .05.
Because these coefficients were consistent in direction and magnitude
across thresholds (differences ≤ 0.4), the proportional-odds assumption
was considered tenable. A visual inspection provided further support for
this conclusion.

```{r}
ggplot(tbl, aes(threshold, estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - 1.96 * std.error,
                    ymax = estimate + 1.96 * std.error), width = 0.1) +
  labs(x = "Response Threshold", y = "Slope for act_pm_c",
       title = "Threshold-specific slopes (proportional-odds check)")
```

### Model specification and estimation

Following diagnostic checks, I prepared to fit the conditional CLMM
predicting ordinal verbal consent-cue ratings from within- and
between-person predictors of encounter-level alcohol consumption
(`drinks_wp`, `drinks_pm_c`) and sexual behavior type (`act_wp`,
`act_pm_c`), controlling for social desirability bias (`bidr_sc`).

#### Evaluating time effects

Each daily survey was distributed between 11:00 AM and 2:00 PM (with a
reminder two hours later) and asked about experiences *since the
previous survey*, yielding roughly 24-hour retrospective reports. The
diary design may have introduced predictable temporal structure (e.g.,
gradual study fatigue). As such, I will evaluate whether to include a
time-related variable as fixed covariate to adjust for potential
design-related patterns that could otherwise inflate between-person
variance or bias fixed-effect estimates. Participants contributed an
average of approximately 3.3 event-day entries over the 28-day diary
study and event spacing was irregular, so this covariate would serve as
an approximate indicator of temporal context.

Specifically, I will test whether day on study (numeric index of the
survey day) accounts for systematic variation in the ordinal outcomes.
To do so, I fit a CLMM predicting ordinal ratings of verbal cue use
(`verb_bins`). Study day (grand-mean centered; `day_c`) was entered as a
fixed predictor, and a random intercept for participant (`nid`)
accounted for repeated measures.

```{r}
d <- d %>%   # grand-mean center study day
  mutate(day_c = day - mean(day, na.rm = TRUE))

m_day <- clmm(verb_bins ~ day_c + (1 | nid), 
              data = d, link = "logit", Hess = TRUE)

anova(clmm_null, m_day)
summary(m_day)
```

A likelihood-ratio test comparing models with and without the day term
indicated that inclusion of `day_c` did not significantly improve model
fit, $\chi^2(1)= 2.28$, $p = .13$. The estimated fixed effect of day was
positive but nonsignificant ($b= 0.019$, $SE = 0.013$, $z = 1.51$),
suggesting no reliable linear trend in verbal cue ratings across days.

Visual inspection of mean ordinal verbal scores by study day showed
minor fluctuations without a consistent upward or downward trajectory.

```{r}
d %>%
  group_by(day) %>%
  summarise(mean_verb = mean(as.numeric(verb_bins), na.rm = TRUE)) %>%
  ggplot(aes(x = day, y = mean_verb)) +
  geom_line() +
  geom_smooth(method = "loess", se = FALSE) +
  labs(y = "Mean ordinal score (verbal consent)", x = "Study day")
```

After confirming that the linear effect of day was not significant, I
examined potential nonlinear change by fitting a model including
orthogonal polynomial terms for study day up to the cubic order using
the poly() function (raw = FALSE). The model was specified as:

```{r}
m_day_poly <- clmm(verb_bins ~ poly(day_c, 3, raw = FALSE) + (1 | nid),
                   data = d, link = "logit", Hess = TRUE)

anova(clmm_null, m_day_poly)
summary(m_day_poly)
```

A likelihood-ratio test comparing the cubic polynomial model to the
unconditional model indicated that inclusion of the polynomial day terms
did not significantly improve model fit, $\chi^2(3) = 2.33$, $p = .51$.
Consistent with this result, none of the individual polynomial
components were significant ($|z|≤ 1.49$, $p ≥ .13$), suggesting no
reliable linear, quadratic, or cubic change in verbal cue ratings across
days.

#### Specifying fixed-effects

*Need to fill this in without being dreadfully repetitive.*
```{r}
fe_formula <- verb_bins ~ drinks_wp + act_wp + 
  drinks_pm_c + act_pm_c + bidr_sc
```

#### Evaluating random slopes

To determine the appropriate random-effects structure, preliminary CLMMs
were fit with and without random slopes for within-person predictors
(`drinks_wp` and `act_wp`).

```{r}
# 1) Random intercept only (full fixed effects in place)
m0 <- clmm(update(fe_formula, . ~ . + (1 | nid)),
           data = d, link = "logit", Hess = TRUE)

# 2) Add random slope for drinks_wp
m1 <- clmm(update(fe_formula, . ~ . + (1 + drinks_wp | nid)),
           data = d, link = "logit", Hess = TRUE)

# 3) Add random slope for act_wp
m2 <- clmm(update(fe_formula, . ~ . + (1 + act_wp | nid)),
           data = d, link = "logit", Hess = TRUE)

# Compare nested models (fixed effects held constant)
anova(m0, m1)
anova(m0, m2)
```

Likelihood-ratio tests indicated that allowing slopes to vary by
participant did not significantly improve model fit for either predictor
(both p \> .10). Accordingly, a random-intercept–only structure was
retained for subsequent analyses to facilitate model convergence and
interpretability.

#### Conditional effects

I fit a conditional CLMM predicting ordinal verbal consent-cue ratings
from within- and between-person predictors of encounter-level alcohol
consumption (`drinks_wp`, `drinks_pm_c`) and sexual behavior type
(`act_wp`, `act_pm_c`), controlling for social desirability bias
(`bidr_sc`).

```{r}
vars_pre <- c(
  "verb_bins", # outcome
  "nid", # participant ID
  "drinks_wp", "drinks_pm_c", # drink count, within- and between-
  "act_wp","act_pm_c", # sexual behavior type, within- and between-
  "bidr_sc" # time-invariant, person-level social desirability bias
)

# create complete case dataframe
d_cc <- d %>%
  dplyr::select(all_of(vars_pre)) %>%
  tidyr::drop_na() # to ensure equal # of obs per model

# refit unconditional model on complete case df
clmm_null_cc <- clmm(verb_bins ~ 1 + (1 | nid),
                     data = d_cc, link = "logit", Hess = TRUE)

m_pre <- clmm( # fit conditional model
  verb_bins ~ 
    drinks_wp + drinks_pm_c + act_wp + act_pm_c + 
  bidr_sc + (1 | nid), data = d_cc,
  link = "logit", Hess = TRUE)
```

I then compared it to the unconditional (random-intercept-only) model.

```{r}
anova(clmm_null_cc, m_pre)
```

A likelihood-ratio test indicated that the inclusion of these
conditional effects significantly improved model fit, 
$\chi^2(5) = 48.19$, $p < .001$, with a corresponding decrease in AIC from
1379.3 to 1341.1. These results suggest that the set of predictors
collectively explained significant between- and within-person
variation in ordinal verbal consent cue use.

#### Examining fixed-effects

```{r}
summary(m_pre)
performance:::r2(m_pre)
```

Within-person differences in alcohol consumption were not associated
with changes in verbal cue use ($b=-0.08$, $p=.40$), but participants
who generally drank more than others before sexual events reported lower
verbal cue use overall ($b=-0.42$, $p=.009$). In contrast, encounters
involving penetrative sexual behavior were associated with markedly
greater verbal consent-cue use both within ($b=1.41$, $p<.001$) and
between ($b=1.46$, $p=.006$) persons. Higher social-desirability scores
also predicted slightly greater reported cue use ($b=0.04$, $p=.005$).
The marginal $R^2$ = 0.14 and conditional $R^2$ = 0.62 indicate that the
model’s fixed effects explained about 14% of total variance, with
substantial additional variation accounted for by individual
differences.

#### Potential Problems and Alternative Strategies

Considering the multiple subscales and items per domain in Aim 1, if
diagnostics reveal predictors are highly correlated, we will prioritize
individual attitudes over perceived peer attitudes and drop the latter
when indicated by $|r|$ ≥ .7 - .8.

If models do not converge in a maximum likelihood framework, they will
be estimated in a Bayesian framework using the brms package in R. 

#### Limitations and Future Directions

The proposed research is not without limitations. First, there were
measures not included in this data collection effort that we intend to
prioritize in the future. Considering this project focuses on active
consent communication, we did not measure passive “no response” signals,
which are commonly reported cues used to infer and communicate consent
among young adults despite risks associated with their ambiguity and
inconsistent associations with internal willingness.2 Also, we did not
assess refusal cues that participants may have used to indicate lack of
willingness, so it is possible that participants were also reflecting on
behaviors they used to assert boundaries when responding to items
assessing verbal and nonverbal consent cues. Other measures not included
that could have strengthened the current study are perceptions of
partner cue usage, because we cannot speak to how participants inferred
consent from their partners without those data, nor can we know whether
the body language, behaviors, or comments that participants interpreted
as cues were intended to signal consent. In future research, these
limitations can be addressed through dyadic data collection, though
recruiting in pairs introduces a different set of constraints with
respect to feasibility and generalizability. Finally, we did not provide
options to select from or ask participants to report specific examples
of the behaviors that they were classifying as verbal or nonverbal cues,
and we did not provide participants with the operational definitions for
these constructs prior to study initiation. 

Specific characteristics of the current study sample and design also may
limit what can be accomplished with the proposed research. Participants
were recruited via Sona Systems and compensated with extra credit in
their introductory psychology class, resulting in a college-student
sample with 85% of participants in their first year. More than 90% of
the sample fell below age 20, while a sample with more proportional
representation of upperclassmen may have yielded a greater number of
partnered sexual events and drinking episodes for analysis. Further, as
these data were collected from college students at a primarily white
institution in the midwestern U.S., the sample is homogeneous in terms
of race (77% White) and sexual orientation (87% heterosexual).
Directions for future research include more stringent criteria that
require eligible study participants be sexually active; prior sexual
experience was not an inclusion criterion for participating in the
current research, as it is common for young adults’ first partnered
sexual experiences to occur during their first semester at
university.[@follingstad2023] 

With respect to study design, the experience sampling method used to
support this project was a relative strength compared to the primarily
cross-sectional literature in this area. Using electronic diaries
reduced the timeframe for retrospective recall, however, survey timing
and framing still required participants to think backwards when
reporting on behaviors and experiences that preceded sexual events
(e.g., pre-encounter affect, drink estimates) on the previous day. Thus,
results examining event-level variables can only be interpreted as
associational and predictive within the concurrent window. Additionally,
asking participants to rate, for example, a sexual partner’s
intoxication level relative to their own at the same timepoint that we
asked them about sexual consent communication may have increased
susceptibility to social desirability bias, potentially leading
participants to respond in ways they thought would be perceived more
positively by the research team, or in ways that reduced their own
cognitive dissonance around possibly taking advantage of an intoxicated
person or being taken advantage of themselves. In future research, these
limitations could be addressed using ecological momentary assessment
with ≥2 surveys per day and/or event-contingent sampling, permitting
assessment of contextual variables pre-encounter and
concurrent/post-encounter variables immediately thereafter. Finally,
there is risk of overfitting to sample-specific nuance, as we are
fitting and evaluating our models in the same sample; techniques for
estimating out-of-sample generalizability were considered but ultimately
omitted from the analytic plan, which prioritizes feasibility.
Directions for future research include goals to replicate our findings
in a separate college sample of new participants, as well as among young
adults in other sociocultural contexts.

## References

::: {#refs}
:::
