---
title: "Data cleaning and screening"
output: 
  html_document:
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{css, echo = FALSE}
pre, code {
  max-height: 500px;
  overflow-y: auto;
  white-space: pre !important; 
  overflow-x: auto
}
```

```{r}
# to detect and warn about conflicts
library(conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

```{r}
# to read in data
library(readr)

path_nids <- "_data/data_process"
path_ready <- "_data/data_ready"

# for data wrangling and cleaning
library(tidyverse)
library(janitor)
library(lubridate)
library(psych)

source(here::here("_scripts/functions.R"))
```

```{r, warning=FALSE}
baseline <- read_csv(here::here(path_nids,"baseline_nid.csv"), 
                     show_col_types = FALSE)

base_sona <- read_csv(here::here(path_nids,"baseline_sona_id.csv"), 
                     show_col_types = FALSE)

data <- read_csv(here::here(path_nids, "diary_nid.csv"), col_types = cols(
  StartDate = col_character(),
  EndDate = col_character(),
  RecordedDate = col_character(),
  AuDB4 = col_character(),
  AuDB4_7_TEXT = col_character(),
  AuDB5 = col_character(),
  AuDB5_6_TEXT = col_character(),
  Q186 = col_character(),
  Q186_8_TEXT = col_character(),
  PSB2 = col_character(),
  Q211 = col_character(),
  PSB15 = col_character(),
  OSBB2 = col_character(),
  OSBB6 = col_character(),
  OSBB10_11 = col_character(),
  SSAB3 = col_character(),
  SSAB6_7 = col_character(),
  Q215 = col_character())) %>%
    clean_names() %>%
    glimpse()
```

## Initial cleaning

Format dates

```{r}
data <- data %>% 
  mutate(start_date = as_datetime(start_date, tz = "America/Chicago"),
         end_date = as_datetime(end_date, tz = "America/Chicago"),
         recorded_date = as_datetime(recorded_date, tz = "America/Chicago"),
         date = date(start_date))
```

Confirm participant identifiers

```{r}
nrow(data) # 6928
length(unique(data$nid)) # 278 - nid derived from sona_id; links baseline to daily surveys
length(unique(data$pid)) # 278 - pid exclusive to daily surveys
```

### Remove blank surveys

Remove 136 completely blank surveys

```{r}
data_empty <- data %>% 
  select(au_db1:e4) 

data_empty_rows <- which(rowSums(is.na(data_empty)) == ncol(data_empty))

length(data_empty_rows)

data <- data %>% 
  filter(!row_number() %in% data_empty_rows)
```

### Identify duplicates

255 duplicates (i.e., more than one survey on same day for one pid)

```{r}
data %>% 
  group_by(pid, date) %>% 
  count() %>% 
   ungroup() %>% 
  filter(n > 1) %>% 
  arrange(desc(n))
```

Check if responses are same on duplicates

```{r}
duplicates <- data %>% 
  group_by(pid, date) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  left_join(data, by = c("pid", "date")) %>% 
  filter(!is.na(n)) 

duplicates %>% 
  distinct() %>% 
  nrow() 

duplicates %>% 
  select(-c(4:20)) %>% 
  distinct() %>% 
  nrow()

```

Remove duplicates that are same on all columns except survey metadata\
Will keep first survey so we have start time of first survey that day

```{r}
duplicate_responses <- which(duplicates %>% 
  select(-c(4:20)) %>% 
  duplicated())

duplicate_ids <- duplicates %>% 
  filter(row_number() %in% duplicate_responses) %>% 
  pull(response_id)

data <- data %>% 
  filter(!response_id %in% duplicate_ids)
```

Left with 197 duplicate surveys

```{r}
data %>% 
  group_by(pid, date) %>% 
  count() %>% 
   ungroup() %>% 
  filter(n > 1) %>% 
  arrange(desc(n))


duplicates <- data %>% 
  group_by(pid, date) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  left_join(data, by = c("pid", "date")) %>% 
  filter(!is.na(n)) 
```

Check for duplicate surveys with no data other than screening questions
for branching

```{r}
data_empty <- duplicates %>% 
  select(au_db2:e4) 

data_empty_rows <- which(rowSums(is.na(data_empty)) == ncol(data_empty))

duplicates %>% 
  filter(row_number() %in% data_empty_rows) %>%
  group_by(nid, date) %>%
  count() %>% 
  arrange(desc(n))

# 1 nid has 2 mostly blank duplicates - do they have another survey that day?
# they do!
duplicates %>% 
  filter(nid == "5140108" & date == "2022-11-16")
```

filter out mostly blank duplicates (i.e., only responded to branching
questions and have more complete survey on same day)

```{r}
response_ids <- duplicates %>% 
  filter(row_number() %in% data_empty_rows) %>% 
  pull(response_id)

data <- data %>% 
  filter(!response_id %in% response_ids)
```

159 duplicates left

```{r}
data %>% 
  group_by(pid, date) %>% 
  count() %>% 
   ungroup() %>% 
  filter(n > 1) %>% 
  arrange(desc(n))


duplicates <- data %>% 
  group_by(pid, date) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  left_join(data, by = c("pid", "date")) %>% 
  filter(!is.na(n)) 
```

Look at surveys taken at midnight to see if date is incorrect due to
being slightly after midnight

```{r}
data %>% 
  filter(hour(start_date) < 2) %>% 
  select(nid, start_date)
```

5/8 Duplicate surveys taken at midnight and later that night and are
missing survey day before- determined midnight survey was for previous
day. Correcting new date variable here.

Note: this date variable lines up roughly to day on study

```{r}
# data |> 
#   filter(nid == "9796568") |> 
#   select(date, response_id, start_date) |> 
#   print(n = Inf)
# 
# data |> 
#   filter(nid == "9740303") |> 
#   select(date, response_id, start_date) |> 
#   print(n = Inf)
# 
# data |> 
#   filter(nid == "6681525") |> 
#   select(date, response_id, start_date) |> 
#   print(n = Inf)
# 
# data |> 
#   filter(nid == "3256800") |> 
#   select(date, response_id, start_date) |> 
#   print(n = Inf)
# 
# data |> 
#   filter(nid == "9328313") |> 
#   select(date, response_id, start_date) |> 
#   print(n = Inf)
# 
# data |> 
#   filter(nid == "2564475") |> 
#   select(date, response_id, start_date) |> 
#   print(n = Inf)
# 
# data |> 
#   filter(nid == "8311815") |> 
#   select(date, response_id, start_date) |> 
#   print(n = Inf)

data <- data |> 
  mutate(date = if_else(response_id == "R_28GUR3cJzRXaTPC", date("2022-11-10"), 
                        date),
         date = if_else(response_id == "R_3lEREAWesaFQSnG", date("2022-10-28"),
                        date),
         date = if_else(response_id == "R_21G38UK4RkS5Ky5", date("2022-10-28"),
                        date),
         date = if_else(response_id == "R_2q1tWYtBw0XbZlM", date("2022-11-05"),
                        date),
         date = if_else(response_id == "R_2c63k6bRzLwkyJT", date("2022-11-14"),
                        date))
```

Look at differences in branching questions among duplicates

```{r}
duplicates <- data %>% 
  group_by(pid, date) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  left_join(data, by = c("pid", "date")) %>% 
  filter(!is.na(n)) 

duplicates_branch_match <- duplicates %>% 
  select(nid, date, au_db1:porn_screen_fu) %>% 
  distinct() %>% 
  group_by(nid, date) %>% 
  count() %>% 
  filter(n == 1) %>% 
  select(-n) %>% 
  left_join(duplicates, by = c("nid", "date"))
```

Look at mental health questions

```{r}
duplicates_branch_match %>% 
  filter(au_db1 == 2 & psb1 == 2 & osbb1 == 2 & ssab1 == 2 & 
                 drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 2) %>% 
  select(nid, date, response_id, start_date, mis1:e4) %>% 
  glimpse()
```

If answers same keep more complete survey then first survey

```{r}
data <- data %>% 
  filter(!response_id %in% c("R_3nl1aO9wz2HoliI", "R_3j0cQeVBMt5s0Y0",
         "R_27UvVfdaNbLQIvA", "R_1roCa7SNA64ll6b", "R_3JjZhfFzAc87KLQ",
         "R_3eb3ZXUu8ma62HW", "R_1goOMvxm2OO37xE", "R_2anOsrKT0rWfpyh",
         "R_2v7x6suXsPA7lCE", "R_3n1q190pGvgnJvW", "R_3Mmsiry2UFC250F",
         "R_1CD8AVjRYbvFOSE", "R_1pGNAcQtKqkDl4Y", "R_1Kp50zwQe1EHJ9A"))
```

If answers differ only by 1-2 items, keep more complete survey or if
both complete keep first survey

```{r}
data <- data %>% 
  filter(!response_id %in% c("R_2e3FNxw2bfXqVhD", "R_3PuSSFXqM2rg75H",
                             "R_21jN639l8fmHhv7", "R_24cL375YzQDqcNs",
                             "R_27InzwQuaYLA91G", "R_2q3iIQW0ju1U0TC",
                             "R_2aQz1zTeseotOnq", "R_aYqSgHFF1qJem9r",
                             "R_1mI5Ldgr8fNIDJn", "R_5mZeozvhOLfQHyF",
                             "R_3G2MmYKXMfOPpw8", "R_12hVYRgqDPJX7KM",
                             "R_V418rtMyfoi5rTb", "R_1zfPXLu96CO1qfv", 
                             "R_1FOH5eJeKZKDLaE", "R_vMHmj0vxM2bn98R",
                             "R_1lt7qIcUoYRj7s2", "R_vAMr55UaumNxt9n",
                             "R_xbVy68nr7ZEHK01", "R_12QL6qI8bCV2YPL",
                             "R_3rZiJxcPpT5dgx4", "R_27PDNNBEMijZqxM",
                             "R_3FJ7693Okhf5xgn", "R_DFXWiS5oKvsSevn",
                             "R_uxoQLKB3iUwZvJ7", "R_5aMdcpvNUaRgepH",
                             "R_1GWuG4iIs1EgIdy", "R_3nUxYS964qGdnyn",
                             "R_1jIj5b2QDOMaSgO", "R_2QFE8b5dMtmd3JL",
                             "R_ZjK1J5cruthIO6R", "R_1kZV0eMBz42WEFW",
                             "R_1FfEfAVl19LMXFw", "R_3gZOFZK2937JZqj"))
```

Look at other matching branch surveys

```{r}
# duplicates_branch_match %>% 
#   filter(au_db1 == 1 & psb1 == 2 & osbb1 == 2 & ssab1 == 2 & 
#                  drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 2) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
# 
# duplicates_branch_match %>% 
#   filter(au_db1 == 1 & psb1 == 1 & osbb1 == 2 & ssab1 == 2 & 
#                  drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 2) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
# 
# duplicates_branch_match %>% 
#   filter(au_db1 == 2 & psb1 == 2 & osbb1 == 1 & ssab1 == 1 & 
#                  drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 2) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
# 
# duplicates_branch_match %>% 
#   filter(au_db1 == 2 & psb1 == 2 & osbb1 == 2 & ssab1 == 1 & 
#                  drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 2) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
# 
# duplicates_branch_match %>% 
#   filter(au_db1 == 2 & psb1 == 2 & osbb1 == 2 & ssab1 == 1 & 
#                  drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 1) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
# 
# duplicates_branch_match %>% 
#   filter(au_db1 == 2 & psb1 == 1 & osbb1 == 2 & ssab1 == 1 & 
#                  drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 1) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
# 
# duplicates_branch_match %>% 
#   filter(au_db1 == 1 & psb1 == 2 & osbb1 == 2 & ssab1 == 2 & 
#                  drugs_screen_fu == 2 & bpa1 == 1 & porn_screen_fu == 2) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
# 
# duplicates_branch_match %>% 
#   filter(au_db1 == 2 & psb1 == 2 & osbb1 == 1 & ssab1 == 2 & 
#                  drugs_screen_fu == 2 & bpa1 == 2 & porn_screen_fu == 1) %>% 
#   select(nid, date, response_id, start_date, au_db2:e4) %>% 
#   view()
```

If answers differ only by 1-2 items (or 3 items if more than one branch
endorsed and completed), keep more complete survey or if both complete
keep first survey

```{r}
data <- data %>% 
  filter(!response_id %in% c("R_1DYDNJfbNuwRivW", "R_3w9vDPa7Aoq1lFD",
                             "R_22Gkq3271dAoUP2", "R_1gTtr4rJocjW0Uy",
                             "R_2uJW6wj5ne3i1wS", "R_3MGuKFZITFIe3NM",
                             "R_3MDR4RbykC162QV", "R_3Gq3wQfAmpYVFit",
                             "R_3Mn9DViM1ZJlQ1Q", "R_3dYXmYQcqJbtXbD",
                             "R_Ww9Ie3dTvQ58QbT", "R_1OVRbIUU6JzciJI",
                             "R_1P1q8xsYV2BQ3KG", "R_DclSImfNg9xma6B",
                             "R_3s17B90hYoNIXpy", "R_3Lb7H5oppSanIB0",
                             "R_Aafzxghrkp8ZNEl"))
```

Look at non-matching branch surveys

```{r}
duplicates %>% 
  filter(!response_id %in% duplicates_branch_match$response_id) %>% 
  select(nid, date, response_id, start_date, au_db1:e4) %>% 
  glimpse()
```

Duplicates by nid

```{r}
# total duplicates
duplicates %>% 
  group_by(nid) %>% 
  count() %>% 
  arrange(desc(n))

# days with duplicates
duplicates %>% 
  group_by(nid, date) %>%
  slice(1) %>% 
  group_by(nid) %>% 
  count() %>% 
  arrange(desc(n))
```

99 duplicates remaining

```{r}
duplicates <- data %>% 
  group_by(pid, date) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  left_join(data, by = c("pid", "date")) %>% 
  filter(!is.na(n)) 

data %>% 
  group_by(pid, date) %>% 
  count() %>% 
   ungroup() %>% 
  filter(n > 1) %>% 
  arrange(desc(n)) %>% 
  print(n = Inf)
```

------------------------------------------------------------------------

Manually inspected remaining duplicates

Check how much time between duplicate surveys

```{r}
duplicates %>% 
  group_by(nid, date) %>% 
  summarize(time_between_dups = max(start_date) - min(start_date)) %>% 
  arrange(time_between_dups) %>% 
  print(n = Inf)
```

Look at duplicates more than 3 hours apart (3600 seconds = 1 hour)

```{r}
duplicates %>% 
  group_by(nid, date) %>% 
  summarize(time_between_dups = max(start_date) - min(start_date)) %>% 
  arrange(time_between_dups) %>% 
  filter(time_between_dups > 10800) %>% 
  print(n = Inf)

duplicates %>% 
  group_by(nid, date) %>% 
  summarize(time_between_dups = max(start_date) - min(start_date)) %>% 
  filter(time_between_dups > 10800) %>% 
  ungroup() %>% 
  left_join(duplicates, by = c("nid", "date")) %>% 
  select(nid, date, response_id, progress, everything()) %>% 
  glimpse()
```

```{r}
data <- data %>% 
  filter(!response_id %in% c("R_117vFnoNfgFOk4D", "R_d0WuHXIYc3QHGyl",
                             "R_2OSU9MciQO5OGhT"))
```

If progress \< 100 and another is 100 remove \< 100 duplicate

```{r}
data <- data %>% 
  filter(!response_id %in% c("R_3PzBQDOn70O3RHN", "R_2VgBsuP6FAg18p7", 
                             "R_2ZQxY4bcYnHkKVT", "R_2ANi9iaiZrRiK7W",
                             "R_zYcdQAFTkd4Wf85", "R_2RUMEtkgHOoKiU9",
                             "R_0HP8PsuElUEqq4N", "R_2DZw0MQSQ4Zv2M6",
                             "R_1Dqyd9EAWF6zkGf", "R_12KJL68mIqMDtil",
                             "R_3szg3OAQjR1LWAS", "R_2t36lXAL1Zxe8VG",
                             "R_ZIBMUETiZP7IX2F", "R_1JDaJzJxE3tFhJM",
                             "R_2s0dOdGygUSWdln", "R_2ysynCwzs7kOGy7",
                             "R_cIyJ7yF1Re01kC5", "R_3MPXwxSXDlhGQaI"))
```

```{r}
duplicates <- data %>% 
  group_by(pid, date) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  left_join(data, by = c("pid", "date")) %>% 
  filter(!is.na(n)) 

data %>% 
  group_by(nid, date) %>% 
  count() %>% 
   ungroup() %>% 
  filter(n > 1) %>% 
  arrange(desc(n)) %>% 
  print(n = Inf)
```

For final ones, keep first survey (no way to determine which one is more
accurate)

```{r}
first_surveys <- duplicates %>% 
  group_by(nid, date) %>% 
  arrange(date) %>% 
  slice(1)


data <- data %>% 
  filter((response_id %in% duplicates$response_id & !response_id %in% first_surveys$response_id) | (response_id %in% data$response_id & !response_id %in% duplicates$response_id))
```

All duplicates resolved

```{r}
duplicates <- data %>% 
  group_by(pid, date) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  left_join(data, by = c("pid", "date")) %>% 
  filter(!is.na(n)) 

data %>% 
  group_by(nid, date) %>% 
  count() %>% 
   ungroup() %>% 
  filter(n > 1) %>% 
  arrange(desc(n)) %>% 
  print(n = Inf)
```

### Check number of surveys per unique participant

```{r}
data %>% 
  group_by(nid) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  print(n = Inf)
```

#### Mean number of surveys completed: 23.5

```{r}
data %>%
  count(nid, name = "n_surveys") %>%
  summarise(
    total_completed = sum(n_surveys),
    total_possible  = n() * 28,
    mean_completed  = mean(n_surveys),
    completion_rate = mean(n_surveys) / 28
  )

# double-check mean # of surveys 
data %>%
  count(nid) %>%
  summarise(avg_surveys_completed = mean(n))

# median, SD, min, max 
data %>%
  count(nid, name = "n_surveys") %>%
  summarise(
    median_surveys = median(n_surveys),
    sd_surveys     = sd(n_surveys),
    min_surveys    = min(n_surveys),
    max_surveys    = max(n_surveys),
    n_participants = n()
  )
```

### Rename variables for consistency

```{r}
d <- rename_variables(data)
```

### Remove unnecessary variables

Either added during cleaning for checks or single value metadata columns

```{r}
d <- d %>% 
  select(-c(r_date, time, rt, rid, ssid, sms, sig,
            recipient_last_name, recipient_first_name, recipient_email,
            external_reference, distribution_channel, user_language)) %>% 
  relocate(nid, date) %>% 
  glimpse()
```

Items from online and solo sexual behavior blocks, which fall beyond the
scope of current project

```{r}
d <- d %>% 
  select(-c(osb2, osb2.txt, osb6:osbb10_11_10_text, osbb15, osbb15_5_text,
            aro_of, pls_of, pss, exp_o, str_o:er4_o, ssb3, ssb3.txt,
            ssb6:er4_s, dub2.txt, med1_d:exp_d, dub3_bf.txt, dub3_du.txt,
            dub3_af.txt, status, q186_8_text, au_l:au_pbs.txt, 
            start_date:end_date, ip_address, recorded_date, 
            location_latitude, location_longitude, group, psb5a, psb6a:str_p.txt, 
            psb7a:psb15_tried_10_text, psb16_10_text:psb16_tried_10_text))
```

## Recoding for inclusion/exclusion

### Daily behavioral screeners

1 = Yes, 0 = No, 3 = Prefer not to say

```{r}
d <- d %>% 
  mutate(across(aub1:pub1, as.numeric)) %>% 
  mutate(across(c(aub1:psb1, osb1:ssb1, pub1),
               ~if_else(.x == 2, 0, .x)))

d <- d %>%  
  mutate(across(c(dub1:bpa),
               ~case_when(. == 2 ~ 0, 
                          . == 4 ~ 3,
                          TRUE ~ .)))

```

Dichotomous variables

1 = Yes, 0 = No

```{r}
d <- d %>%
  mutate(across(
    c(au_pg, au_dg, prec, er1_p, er2_p, er3_p, er4_p),
    ~ case_when(
      . == "1" ~ 1, # Yes
      . == "2" ~ 0, # No
      TRUE ~ as.numeric(.)
    )
  ))
```

### Type of sexual activity

`s_act`: What activities occurred? (check all that apply)

Response options:

-   Kissing = 1;
-   Touching = 2;
-   Digital penetration (i.e., w/ fingers) = 3;
-   Oral sex = 4;
-   Vaginal sex = 5;
-   Anal sex = 6;
-   I tried but no sexual contact happened = 7

```{r}
d %>% select(s_act) %>% map(tabyl)

d <- d %>%
  mutate(s_act = as.character(s_act),
         s_act = case_when(
           s_act %in% c("1", "1,2", "2") ~ "Contact",
           s_act == "7" ~ "Attempt",
           grepl("3", s_act) | grepl("4", s_act) | 
             grepl("5", s_act) | grepl("6", s_act) ~ "Penetration",
           TRUE ~ as.character(s_act)))

d %>% select(s_act) %>% map(tabyl)
```

`psb_di` 0 = Attempted sexual contact; 1 = Contact and/or penetration

```{r}
d$psb_di <- d$s_act

d <- d %>%   
mutate(across(c(psb_di),
                ~dplyr::recode(.,
                               "Attempt" = 0,
                               "Contact" = 1,
                               "Penetration" = 1)))

d %>% select(psb_di) %>% map(tabyl)
```

### Subset diary data

To only include obs for which the partnered sexual behavior daily
screener (psb1) was endorsed AND sexual contact (psb_di) was reported.

```{r}
data_subset <- d %>%
  filter(psb1 == 1 & psb_di == 1)

data_subset %>%
  summarise(
    n_obs = n(),
    n_participants = n_distinct(nid)
  )
```

## Descriptive checks

### Missing diary data

```{r}
d <- data_subset %>%
  select(nid, date, response_id, s_act, aro_pf, pls_pf,
         verb, enon, inon, prec, fm_r, drink, intox, p_tox, bpa)

colSums(is.na(d))[colSums(is.na(d)) > 0]
```


*Outcomes = Verbal and Nonverbal sexual consent cue use*.

### Verbal consent

To what extent did you do the following during the sexual activity?

`verb`: Directly asked your partner or communicated verbally with your
partner about consent

```{r}
psych::describe(d$verb)
```

```{r}
d %>% select(verb) %>% map(tabyl)

d %>%
  pull(verb) %>%
  hist(main = "Histogram of Verbal Cues",
       xlab = "Verbal Self-Report",
       col = "lightblue",
       breaks = 10)
```


```{r, eval=FALSE}
#### Verbal - Binned for CLMMs

d <- d %>%
  mutate(
    verb_bins = cut(
      verb,
      breaks = c(-Inf, 0, 3, 6, 9, Inf),
      labels = c("Not at all", "A little", 
                 "Somewhat", "Quite a bit", "Very much"),
      right = TRUE,
      include.lowest = TRUE),
    verb_bins = ordered(verb_bins,
                        levels = c("Not at all", "A little", 
                                   "Somewhat", "Quite a bit", "Very much")))

# Verify
is.ordered(d$verb_bins)
table(d$verb_bins, useNA = "ifany")
```

### Explicit nonverbal

To what extent did you do the following during the sexual activity?

`enon`: Used direct non-verbal signals to communicate with your partner
about consent for the sexual activity

```{r}
psych::describe(d$enon)
```

```{r}
d %>% select(enon) %>% map(tabyl)

d %>%
  pull(enon) %>%
  hist(main = "Histogram of Direct Nonverbal",
       xlab = "Direct Nonverbal Self-Report",
       col = "lightblue",
       breaks = 10)
```

### Implicit nonverbal

To what extent did you do the following during the sexual activity?

`inon_pc`: Used subtle signals to communicate with your partner about
consent for the sexual activity

```{r}
psych::describe(d$inon)
```

```{r}
d %>% select(inon) %>% map(tabyl)

d %>%
  pull(inon) %>%
  hist(main = "Histogram of Indirect Nonverbal",
       xlab = "Indirect Nonverbal Self-Report",
       col = "lightblue",
       breaks = 10)
```

#### Correlation: Explicit - Implicit

```{r}
round(cor(d$enon, d$inon, use = "pairwise.complete.obs"), 2)
```

The correlation between the `direct` and `indirect nonverbal` consent
cue items was *r* = .73, indicating a strong positive relationship that
suggests these items reflect a shared underlying construct. For
parsimony and theoretical clarity, these items will be averaged to
create a continuous composite score representing overall nonverbal
consent communication.

### Nonverbal composite

*Re-calculate post-imputation*

```{r}
d <- d %>%
  rowwise() %>%  # calculate mean score
  mutate(nvrb = mean(c(enon,inon)))

psych::describe(d$nvrb)
```

#### Correlation: Verbal - Nonverbal

```{r}
# Check correlation between verbal and nonverbal 
round(cor(d$verb, d$nvrb, use = "pairwise.complete.obs"), 2)
```

```{r, eval=FALSE}
#### Nonverbal - Binned for CLMMs
# Create ordered bins for nonverbal consent cues
d <- d %>%
  mutate(
    nonvb_bins = cut(
      nvrb,
      breaks = c(-Inf, 0, 3, 6, 9, Inf),
      labels = c("Not at all", "A little", 
                 "Somewhat", "Quite a bit", "Very much"),
      right = TRUE,
      include.lowest = TRUE
    ),
    nonvb_bins = ordered(nonvb_bins,
                         levels = c("Not at all", "A little", 
                                    "Somewhat", "Quite a bit", "Very much")))
# Verify
is.ordered(d$nonvb_bins)
table(d$nonvb_bins, useNA = "ifany")
```

### Overview L1 predictors

#### Sexual precedent

`prec`: Had you ever had sexual contact with this person in the past?

No = 0; Yes = 1

```{r}
d %>% select(prec) %>% map(tabyl)
```

#### Partner familiarity (raw)

`fm_r`: How did you know this other person?

I didn’t, we just met that day = 1; They are a friend = 2; They are an
acquaintance = 3;\
They are my current or past dating/intimate partner = 4

```{r}
d %>% select(fm_r) %>% map(tabyl)
```

#### Participant drink count

`drink`: How many drinks did you consume prior to or during this
activity?

Response options: 0 = 1; 1-2 = 2, 3-4 = 3, 5-6 = 4, 7-9 = 5, 10+ = 6

```{r}
d %>% select(drink) %>% glimpse()

d %>%
  pull(drink) %>%
  hist(main = "Histogram of Participant Drinks",
       xlab = "Self-Reported Drink Count",
       col = "lightblue",
       breaks = 10)
```

-   Recode `psb_alc` from `drink` such that zero drinks = 0 and \>0
    drinks = 1

```{r}
d <- d %>%
  mutate(psb_alc = if_else(drink == 1, 0, if_else(drink > 1, 1, drink)))

d %>% select(psb_alc) %>% map(tabyl)
```

Participants endorsed `r nrow(subset(d, psb_alc == 1))` partnered sexual
events following alcohol consumption.

#### Self-reported intoxication

`intox`: How intoxicated did you feel right before or during this sexual
activity?

Least intoxicated ever felt in life (0) - Most intoxication ever felt in
life (100) VAS

```{r}
d %>%
  filter(psb_alc == 1) %>%
  pull(intox) %>%
  hist(main = "Histogram of Participant Intoxication",
       xlab = "Self-Reported Intoxication",
       col = "lightblue",
       breaks = 10)
```

#### Perceived partner intoxication

`p_tox`: How intoxicated do you think the other person felt right before
or during this sexual activity?

Least intoxicated ever felt in life (0) - Most intoxication ever felt in
life (100) VAS

```{r}
d %>%
  filter(p_tox > 0) %>%
  pull(p_tox) %>%
  hist(main = "Histogram of Partner Intoxication",
       xlab = "Perceived Partner Intoxication",
       col = "lightblue",
       breaks = 10)
```

#### Drinking environments

Bar-or-party attendance pre-encounter

-   NOTE: n = 1 nid (9128298) endorsed "Prefer not to respond" for bpa
-   treat as missing and impute

```{r}
d %>% select(bpa) %>% map(tabyl)

d <- d %>%
  mutate(bpa = ifelse(bpa == 3, NA, bpa))
```

#### Sexual feelings

*Used for measurement validity assessment, not modeled in confirmatory
analyses.*

To what extent did you feel the following during the sexual activity?

-   `aro_pf`: I felt aroused.

```{r}
d %>% select(aro_pf) %>% map(tabyl)
```

-   `pls_pf`: I felt pleasure

```{r}
d %>% select(pls_pf) %>% map(tabyl)
```

-   **Correlation** between arousal and pleasure

```{r}
round(cor(d$aro_pf, d$pls_pf, use = "pairwise.complete.obs"), 3)
```

-   **Composite** for encounter-level sexual feelings

```{r}
d <- d %>%
  rowwise() %>%  # calculate mean score
  mutate(psb_feel = mean(c(pls_pf,aro_pf)))

psych::describe(d$psb_feel)
```

## Baseline

Replace sona_id with new ID (`nid`) to ensure data remains
de-identified.

```{r}
# names(base_sona)

identical(baseline$response_id, base_sona$response_id)

base_sona_updated <- base_sona %>%
  left_join(
    baseline %>% select(response_id, nid),
    by = "response_id"
  )

baseline <- base_sona_updated %>% 
  select(-c(sona_id, sona_id_confirm))

sum(is.na(baseline$nid))
```

Subset baseline to only include obs for which the same participant ID
(nid) also appears in the diary data.

```{r}
baseline_subset <- baseline %>%
  semi_join(data, by = "nid")

# Check that it worked
all(baseline_subset$nid %in% data$nid)
```

Further subset baseline to only include obs for participants who
reported partnered sexual behavior (psb1) involving sexual contact
(psb_di) at least once during the daily diary portion of the study by
removing rows from baseline_subset where that nid does not appear in
data_subset (i.e., remove participants from baseline with zero psb1 == 1
& psb_di == 1 events).

```{r}
base_ana <- baseline_subset %>%
  semi_join(data_subset, by = "nid")
```

Remove unnecessary variables

```{r}
b <- base_ana %>% 
  select(-c(smu_1:sm_txt, sx_txt, re_txt, ders1:ders18, sd_txt, po_txt, mu_txt,
            sss1:sss23, cpa1:v_ar6, apb1:apb6, phq8_1:phq8_fu, gad7_1:gad7_7, 
            pcl_txt:pcl_20, sx_trade1:porn2, au_ven, pcl, sc0, tb2, pcl_s:hyp_c, 
            ar_txt, v_uc1:v_ar6, v_num, v_rev, v_poly, sss_tf, asrs_1:asrs_6, 
            ra_txt, cud_6m:cud8, aud_s:rma_s, s_life))
```

### Check for missing data

NOTE: `aud2` is missing 4 obs 
NOTE: check variable coding for `race`

```{r}
colSums(is.na(b))[colSums(is.na(b)) > 0] # aud2 missing 4 obs
```

## Overview L2 predictors

### Social desirability bias

-   NOTE: items 1, 3, 5, and 8 are already reverse-coded
-   NOTE: items 9, 11, 12 and 13 are already reverse-coded

```{r}
## BIDR (a = .73) 
# (1-8) 1 = totally disagree, 8 = totally agree
#  Higher scores reflect greater social desirability bias

### Total Scale
b <- b %>%
  rowwise() %>%
  mutate(bidr = sum(bidr_1,bidr_2,bidr_3,bidr_4,bidr_5,
                    bidr_6,bidr_7,bidr_8,bidr_9,bidr_10,
                    bidr_11,bidr_12,bidr_13,bidr_14,bidr_15,bidr_16))
```

### Rape myth rejection

```{r}
## IRMA (a = .80)
# 5 = strongly disagree, 4 = Somewhat..., 3 = Neither disagree nor agree, 2 = Somewhat..., 1 = strongly agree 
# Item values summed for total score, with higher scores reflecting greater rejection of rape myths

b <- b %>%
  rowwise() %>%
  mutate(irma = sum(irma1,irma2,irma3,irma4,irma5,irma6,irma7))
```

### Alcohol and sexual coercion - Personal attitudes

NOTE: check reverse-coding

```{r}
b <- b %>%
  rowwise() %>%
  mutate(asco = sum(asc1, asc2, asc3, asc4, coer1, coer2, 
                    coer3, coer4, coer5, coer6))

b <- b %>%
  rowwise() %>%
  mutate(coer = sum(coer1,coer2,coer3,coer4,coer5,coer6))
```


### Perceived peer attitudes - Alcohol and sexual coercion

```{r}
b <- b %>%
  rowwise() %>%
  mutate(ppat = sum(pa1, pa2, pa3, pa4, pa5, pa6))
```

## Save updated dataframe

```{r}
b |> 
  write_csv(here::here(path_ready, "base_tst.csv"))
```

```{r}
data_subset |> 
  write_csv(here::here(path_ready, "diary_tst.csv"))
```
